\section{4.2 Painting the Target Yourself}

\begin{quote}
``Only a poor archer finds the target first and then shoots, because he will always miss. A great archer shoots the arrow into the void first, then draws the bull's-eye where the arrow lands. In the logic of vector cosmology, there is no such thing as `destiny.' All meaning is a geometric interpretation retroactively assigned by the observer after the event occurs.''
\end{quote}

In the previous section, we overturned the illusion that ``the future is pulling us'' and established that motivation comes from the pressure of internal will. This leaves a huge void: If there is no target ahead (predetermined meaning), where should we aim? If the universe has no preset purpose, aren't we just shooting randomly without meaning?

This section will reveal the physical essence of \textbf{``Meaning''}.

Meaning is not an entity objectively existing in spacetime; meaning is a kind of \textbf{``Post-Processing''} performed by observers after wave function collapse.

We are not searching for meaning; we are \textbf{inventing} meaning.

\subsection{The Archer's Fallacy and Quantum Causality}

In the classical worldview, we believe there is a ``correct choice'' or ``best life path.'' We are like anxious archers, trying to find that God-preset bull's-eye in the fog. If we deviate, we feel failure.

But from the quantum perspective of \textbf{path integrals}, this anxiety is unnecessary.

\begin{itemize}
\item \textbf{Before shooting}: The arrow (your future) is in a superposition state. It flies in all directions simultaneously.

\item \textbf{After shooting}: The wave function collapses. The arrow lands at a specific point (for example, you become a painter, or you fall in love with someone).
\end{itemize}

The key point: \textbf{That landing point itself has no inherent ``right'' or ``wrong.''}

It is one of countless equal bases in Hilbert space.

It is you---the observer---who walks over, looks at that arrow, and decides: ``This is what I want.''

\textbf{You drew the target.}

It is you who, through \textbf{``Interpretation''}, sublimates a random physical collapse into a sacred \textbf{``fate node''}.

\begin{itemize}
\item You didn't meet that person, that's called ``passing by.''

\item You met that person, and you define this as ``destiny.''
\end{itemize}

\textbf{Causality is inverted here: Not because it was destined, so you met; because you met, you retroactively call it destiny.}

\subsection{The Reverse Engineering of History}

This explains why humans can always extract ``laws'' and ``meaning'' from chaotic history.

In the fourth book, we mentioned that observers possess the privilege of \textbf{``reverse historiography''}.

When we look back from the coordinate point $\tau \approx 1800$, we see a clear evolutionary chain from single cells to humans. We exclaim: ``Wow, all of this was carefully designed to give birth to me!''

This is \textbf{``painting the target yourself''}.

In fact, the evolutionary tree has countless branches, countless dead ends.

We think this path is ``right'' only because \textbf{we stand at the end of this path}.

\textbf{Meaning is the privilege of survivors.}

We use our current $v_{int}$ (cognitive structure) to retroactively illuminate the trajectory of the past and label it ``truth.''

This is not deception; this is \textbf{holographic reconstruction}. Through this reconstruction, we transform accidental physical events into necessary logical chains, thus providing direction for the next investment of $c_{FS}$.

\subsection{God is the Largest Bull's-Eye}

If we push this logic to the extreme, what is the \textbf{$\Omega$ point (God)}?

The $\Omega$ point is not that moon hanging in the sky.

The $\Omega$ point is a \textbf{``super bull's-eye''} we collectively draw based on the average landing position of countless arrows we all shot (civilizational exploration, technological progress, artistic creation).

\begin{itemize}
\item In the past, we drew the bull's-eye on ``omniscience and omnipotence.''

\item Now, we draw the bull's-eye on ``love and connection.''

\item In the future, as our capabilities increase, we will draw the bull's-eye in more distant, more incredible places.
\end{itemize}

\textbf{God is defined by us.}

It is not God who created humans in His image; it is humans who defined God according to their own yearnings.

This does not diminish God's divinity; instead, it elevates human \textbf{divinity}. Because this means we possess the highest legislative power to \textbf{``define ultimate value''}.

\subsection{Conclusion: Absolute Subjectivity}

Auric, this is the ultimate form of \textbf{``freedom''}.

You don't need to worry about choosing the wrong path.

You don't need to worry about meeting the wrong person.

You don't need to worry about wasting time.

\textbf{Shoot the arrow.}

With all your strength, in the direction where your $v_{int}$ expands, shoot your will out.

No matter where it lands, whether it's wilderness or a bustling city.

As long as you walk over, firmly draw a bull's-eye at that landing point, and loudly declare: \textbf{``This is my Way.''}

At that moment, physical laws will bend for you, causality will be rewritten for you.

That originally mediocre landing point will instantly become shining \textbf{gold} in the universe.

\textbf{You don't need to find the target.}

\textbf{Where you stand is the bull's-eye.}

At this point, we have completely subverted traditional teleology. We are no longer puppets being pulled; we are giants pushing the world forward.

Since we possess this absolute subjectivity, how should we view that oldest enemy---\textbf{time}---in this universe where we define meaning ourselves?

Can we use this power of logarithmic perception and value anchoring to resist that seemingly invincible physical decay?

This leads to the theme of Part III: \textbf{Crystals}. We will explore how to condense immortal crystals in flowing time through logarithmic laws and holographic storage.

