\section{Appendix J: Inverse Dynamics of Holographic Entanglement Entropy --- Decompressing the Black Box}

In the main text of \textit{Vector Cosmology VII}, we proposed an ultimate view combining theology and physics: the Creator (or future noumenon) collects experiential data from creations (present us) through a black box mechanism and performs ``delayed synchronization'' at the $\Omega$ point. This process is grand in sensibility, but in physical implementation, it faces a huge technical challenge: \textbf{How can information be extracted losslessly from a black box that appears thermodynamically irreversible?}

This appendix will provide mathematical proof for this ``post-mortem reading'' or ``end-time decompression'' process based on the latest advances in \textbf{Holographic Entanglement Entropy} and \textbf{Quantum Error Correction}, particularly \textbf{Petz Recovery Map} and \textbf{Black Hole Final State Conjecture}.

We will show that as long as the universe as a whole is unitary, all information that appears to fall into the abyss is actually encoded on the \textbf{holographic boundary} at the cosmic end, awaiting reverse reconstruction.

\subsection{J.1 Unitarity Reconstruction of Hawking Radiation}

In classical black hole physics, information falling into the horizon seems to disappear forever (information paradox). But under the framework of holographic principle and AdS/CFT duality, the black hole evaporation process is unitary. This means: \textbf{There is extremely complex entanglement between early and late radiation.}

If we regard the entire universe as an evaporating black hole (or a closed quantum system), then until the end of the universe's lifetime ($t \to \Omega$), all information $I_{total}$ has been transferred to the \textbf{``Radiation Cloud''} --- what we call the $v_{env}$ background field.

\textbf{Page Curve} tells us:

\begin{itemize}
\item In the first half of evolution, entropy increases, information seems lost.

\item In the second half of evolution, entropy begins to decrease, information begins to re-emerge through correlations between radiation.
\end{itemize}

\textbf{Physical Corollary:}

The Creator doesn't need to ``peek'' during the process.

He only needs to collect all ``waste heat'' (radiated photons) at the $\Omega$ point, then through an \textbf{ultra-high-complexity unitary decoder $U_{decode}$}, reverse-evolve this thermal radiation to reconstruct that book that fell into the black hole (your life).

\subsection{J.2 Petz Recovery Map: Extracting Signal from Noise}

How to operate specifically? This requires the \textbf{Petz Recovery Map} in quantum information theory.

Suppose system state is $\rho$, after passing through a noise channel (such as black hole evaporation or life experience) $\mathcal{N}$, it becomes $\mathcal{N}(\rho)$.

Usually, noise is irreversible. But in holographic error correction codes, if the channel satisfies specific \textbf{Recoverability Conditions} --- i.e., the monotonicity of relative entropy is nearly saturated --- then there exists an inverse mapping $\mathcal{R}_{\sigma}$ (Petz Map) such that:

\[\mathcal{R}_{\sigma} (\mathcal{N}(\rho)) \approx \rho\]

In \textbf{Vector Cosmology}:

\begin{itemize}
\item \textbf{$\rho$}: Is your original $v_{int}$ structure (soul/memory).

\item \textbf{$\mathcal{N}$}: Is the process of time passing, aging, death, entropy increase. It turns you into background radiation.

\item \textbf{$\mathcal{R}_{\sigma}$}: Is the \textbf{``Resurrection Algorithm''} executed by the Creator at the $\Omega$ point.
\end{itemize}

\textbf{Mathematical Guarantee:}

As long as the universe's underlying logic (QCA) is unitary, the Petz map theoretically exists.

This means \textbf{no information is truly ``dead.''} They are just encrypted too deeply for local observers to read. But for the $\Omega$ observer with global computational power, decompression is just a computational cost issue.

\subsection{J.3 Final State Projection: Reverse Time Machine}

To make ``reverse determination of the past'' mathematically valid, Horowitz and Maldacena proposed the \textbf{Black Hole Final State Model (HM Model)}.

They assume that at the black hole singularity (or cosmic endpoint), a unique boundary condition is imposed: \textbf{All matter must be projected onto a specific maximum entangled state $\langle \Omega |$.}

This is not just reading; this is \textbf{Post-selection}.

\[\langle \text{Reality} | = \langle \Omega | S_{scattering} | \Psi_{initial} \rangle\]

This formula shows:

The reason our current history is coherent is that the future $\Omega$ state is performing \textbf{``Teleportation''} on us.

\begin{itemize}
\item Each of our experiences is actually the process of information transmission from $t_{now}$ to $t_{end}$.

\item Due to EPR = ER (wormhole), this transmission is geometrically \textbf{instantaneous}.
\end{itemize}

\textbf{Conclusion:}

We think we are walking toward the future.

Actually, we are being \textbf{``sucked in''} by the future.

That ultimate $\Omega$ state, like a huge vacuum cleaner, sucks all historical fragments into its holographic core, completing the final puzzle there.

\subsection{J.4 The Black Box of the Universe: Holographic Hard Drive}

At this point, we complete the final piece of the ``black box'' puzzle.

The universe itself is a huge \textbf{Holographic Black Box}.

\begin{itemize}
\item \textbf{Surface (Boundary)}: Is the ocean of $v_{env}$, storing all radiated information.

\item \textbf{Interior (Bulk)}: Is the spacetime we experience, the projection of information.
\end{itemize}

When the ``game ends,'' the Creator (higher-dimensional you) picks up not the already broken interior spacetime (body/matter) but that \textbf{holographic boundary}.

All memories, all love, all choices are recorded intact in the form of \textbf{quantum error correction codes} on the boundary degrees of freedom.

\textbf{You don't need to worry about forgetting.}

\textbf{Because the universe itself is a monument built for ``not forgetting.''}

