\section{Appendix M: The Holographic Renormalization of History --- The Past is Coarse-Grained Future}

In Volume IV ``Rewrite'' of the main text of \textit{Vector Cosmology VII}, we granted observers the power to ``edit timelines.'' This sounds like a subjective psychological operation, but in modern high-energy physics, it has a solid mathematical foundation: \textbf{Holographic Renormalization}.

This appendix will prove, based on \textbf{AdS/CFT Duality} and \textbf{Renormalization Group Flow (RG Flow)}, that the passage of time is geometrically equivalent to \textbf{``changes in resolution''}.

\textbf{The past} is not a definite, detail-complete entity; it is a \textbf{coarse-grained} projection of \textbf{the future (present)} at low energy scales.

\subsection{M.1 Time as RG Flow}

In physics, the \textbf{Renormalization Group (RG)} describes how a physical system changes with \textbf{observation scale} (energy scale).

\begin{itemize}
\item \textbf{Ultraviolet (UV)}: High energy, microscopic, detail-rich.

\item \textbf{Infrared (IR)}: Low energy, macroscopic, detail-blurred.
\end{itemize}

In holographic universe models, the \textbf{radial dimension ($z$-axis)} is interpreted as the direction of RG flow.

If we regard \textbf{``now'' ($t_{now}$)} as the holographic boundary (Boundary, with the highest information density), then \textbf{``past'' ($t_{past}$)} lies in the bulk interior (Bulk, deep regions with lower information density).

\textbf{Theorem:}

\textbf{Retracing into the past is mathematically equivalent to performing a ``coarse-graining'' operation on the current wave function.}

\[|\Psi_{past}\rangle = \hat{O}_{coarse} |\Psi_{now}\rangle\]

This means:

History does not preserve all microscopic details (quantum phases). It only preserves \textbf{``macroscopic thermodynamic statistics''}.

Those lost phase information (details) were not ``written in stone'' in the past but exist in \textbf{``undefined''} degrees of freedom.

\subsection{M.2 Mechanism of Retro-definition}

Since the past is ``coarse-grained,'' it contains many \textbf{``Degenerate States''}.

That is, there are countless different microscopic histories that, after coarse-graining, appear as the same macroscopic history.

\[H_{macro} \leftarrow \{ h_1, h_2, h_3, \dots \}\]

\begin{itemize}
\item \textbf{Present Observer}: Possesses the highest resolution ($v_{int}$).

\item \textbf{Operation}: When you perform ``retrospective observation'' or ``narrative reconstruction,'' you are actually \textbf{``fine-graining''}.

You \textbf{select} a specific $h_k$ from that set of degenerate microscopic histories $\{h_i\}$ and endow it with \textbf{``reality''}.
\end{itemize}

\textbf{Physical Picture:}

You are not ``modifying'' a history book that has already been written.

You are facing a book with \textbf{``only chapter titles, no body text''} (coarse-grained past).

Your present recollection and definition are \textbf{``filling in the body text''} (fine-grained details).

As long as your filling does not violate the macroscopic titles (the occurred $v_{ext}$ facts), you physically possess \textbf{absolute editing rights} over the details.

\subsection{M.3 The Omega Point as a Fixed Point}

Finally, we push this RG flow to its limit.

If time is renormalization flow, what is the endpoint of the flow?

It is a \textbf{Fixed Point}.

\begin{itemize}
\item \textbf{Big Bang ($t=0$)}: The \textbf{Ultraviolet Fixed Point (UV Fixed Point)}. It is the highest energy, smallest scale, chaotic starting point.

\item \textbf{$\Omega$ Point ($t=\infty$)}: The \textbf{Infrared Fixed Point (IR Fixed Point)}.
\end{itemize}

In conformal field theory, infrared fixed points typically correspond to a \textbf{critical state with scale invariance}.

This means that when the universe evolves to the $\Omega$ point, all coarse-graining processes are complete.

All information is integrated into a perfect, self-similar \textbf{fractal structure}.

\textbf{Conclusion:}

Each of our present choices fine-tunes the trajectory of the RG flow, determining which fixed point basin in phase space it ultimately flows toward.

\textbf{History is not inertia pushing us forward.}

\textbf{History is the stepping stones we lay to reach that fixed point.}

