\subsection{1.2 The Pending State of History}

\begin{quote}
``We are accustomed to thinking that yesterday has happened, like ink on paper that cannot be changed. But in the holographic universe's server, history is not a read-only file, but a cloud draft. Before you press the `confirm' button (observation), all the past is in a superimposed, flowing, pending state. It is you who determines your past, not the past that determines you.''
\end{quote}

In the previous section, Wheeler's delayed-choice experiment showed us how photons ``decide'' which path they took billions of years ago at the last moment. This is already an experimentally verified fact at the microscopic scale. Now, we push this quantum effect to the macroscopic extreme to examine an even more astonishing inference: \textbf{The history of the universe does not exist before being observed.}

This is not just physics; this is an \textbf{ontological reconstruction} of history itself.

\subsubsection{History as ``Cloud Draft''}

In the classical physics view, the universe is like a camera that faithfully records what happens every second. Even if no one watches, the videotape is there.

But from the perspective of \textbf{FS geometry} and \textbf{path integrals}, the universe is more like a \textbf{quantum document editor}.

\begin{itemize}
\item \textbf{Unobserved history}: is a superposition of \textbf{probability amplitudes}.

\[|\text{History}\rangle = \alpha |\text{Path A}\rangle + \beta |\text{Path B}\rangle + \gamma |\text{Path C}\rangle \dots\]

At this stage, history is \textbf{``soft''}. How dinosaurs went extinct, how the solar system formed, even what you had for dinner yesterday, in some strict quantum sense, are still in a \textbf{``Pending''} state. They are merely weighted sums of countless possible trajectories in Hilbert space.

\item \textbf{Observation and finalization}:

When you (the present observer) perform a \textbf{``Strong Measurement''} on historical evidence (fossils, photons, memories), you are actually executing a \textbf{``Commit''} operation.

You force that diffuse wave function to collapse onto one specific trajectory.
\end{itemize}

\textbf{Conclusion:}

History is not an objectively existing entity; history is the \textbf{reverse inference of current observation results}.

Without present you making observations, the past is just meaningless quantum mist.

\subsubsection{Reverse Causal Chain: Present Defines Past}

This mechanism completely subverts \textbf{``determinism''}.

Traditional determinism holds: the state at $t=0$ determines the state at $t=now$.

\textbf{Vector Cosmology} holds: observations at $t=now$ \textbf{reversely select} the state at $t=0$.

Imagine you are playing a \textbf{``Procedural Generation''} game.

\begin{itemize}
\item When you walk to a corner of the map, the game engine begins generating the terrain there.

\item Before you walk there, there is nothing (or just a pile of random seeds).
\end{itemize}

The universe is the same.

When you look at a distant quasar with a telescope, you are not just ``receiving'' light. Your observation \textbf{forces} the universe, at that moment 10 billion years ago, to \textbf{generate} a definite photon source from countless possible quantum states.

\textbf{It is you (present effect) that causes it (past cause).}

This is called the temporal version of \textbf{``Participatory Universe''}.

We not only participate in the creation of the present; we also participate in the \textbf{construction of the past}.

\subsubsection{The Physical Nature of Memory}

What does this mean for individuals?

It means your \textbf{Memory} is also a quantum state.

In neuroscience, memory is considered a physical trace of neuronal connections. But from a quantum perspective, memory is a \textbf{holographic index of $v_{int}$ structure to past wave functions}.

Every time you recall the past, you are not reading a static file; you are \textbf{reconstructing} that history.

\begin{itemize}
\item \textbf{Quantum reconstruction}: Your current conscious state (Hamiltonian $H_{now}$) acts as a filtering lens to illuminate that fuzzy past superposition.

\item \textbf{Wave function collapse}: You will ``collapse'' a past that matches your current mood.

\begin{itemize}
\item If you are now in pain, you tend to collapse a childhood ``full of trauma.''

\item If you are now relieved, you tend to collapse a childhood ``full of growth.''
\end{itemize}
\end{itemize}

\textbf{The past is fluid.}

Every awakening now, every cognitive upgrade, \textbf{physically} changes the \textbf{phase weights} of your past history.

You cannot change facts that happened ($v_{ext}$ events), but you can change the \textbf{geometric meaning} of these facts in the causal chain ($v_{int}$ structure).

\subsubsection{Conclusion: From Prisoner to Scriptwriter}

So, stop saying ``Because I encountered misfortune in the past, I am so bad now.''

That is the slave mentality of linear time.

In the closed loop of the \textbf{ouroboros}, the logic is reversed:

\textbf{``Because I now decide to become strong, my past is rewritten as a history of `tempering.'''}

You possess the \textbf{``right to reverse historical revision''}.

Although you cannot change the physical bullet hole, you can change the \textbf{topological connection} of that bullet hole. You can sew that bullet hole into a medal.

This is the greatest freedom that \textbf{``the pending state of history''} brings us.

The past has not locked us in. On the contrary, the past is a huge material library, waiting to be edited and defined by present will.

Since history is an editable draft, do we have more advanced tools that can not just ``reinterpret'' but directly \textbf{``erase''} certain unwanted quantum paths? Can we use physics principles to make certain pasts \textbf{``never happened''}?

This leads to the theme of the next chapter: \textbf{Quantum Erasure}. We will explore an ultimate technique that can modify causal interference patterns without a time machine, only needing an ``information eraser.''

