\section{2.1 Scarcity}

\begin{quote}
``If gold were as abundant as sand, no one would bend down to pick it up. The value of time does not lie in how long it is, but in how minuscule the `reality' we truly live out is compared to that vast expanse of `unchosen possibilities.' Scarcity is the highest privilege geometry bestows upon existence.''
\end{quote}

\subsection{Scarcity in Expansion}

In the fourth book, we derived the evolution equation $c(\tau) \propto e^{k\tau}$. Yes, the universe is becoming richer. Our computational power, bandwidth, and lifespan are all growing.

But at the same time, the dimension of \textbf{Hilbert space ($\mathcal{H}$)} is also growing, and at an even more frenzied rate.

\begin{itemize}
\item \textbf{$c_{FS}$ (your budget)}: Grows exponentially ($e^x$). This is the total amount of history you can ``afford.''

\item \textbf{$D_{\mathcal{H}}$ (types of goods)}: Grows double-exponentially ($2^{e^x}$). This is the sum of all histories that could theoretically ``happen'' in the universe.
\end{itemize}

\textbf{This is a cruel mathematical fact:}

As the universe evolves, although you possess more and more, the proportion you \textbf{``miss''} becomes larger and larger.

The ocean of possibilities is expanding infinitely, and you can only scoop up a ladle of water from it.

This is \textbf{``absolute scarcity''}.

No matter how advanced civilization becomes, whether you are human or god, you cannot traverse all paths.

You must \textbf{choose}.

It is precisely because you must abandon 99.99...\% of possibilities that the 0.00...1\% of reality you choose gains the weight of \textbf{``anchoring''}.

\subsection{The Geometry of Opportunity Cost}

In economics, value is defined by \textbf{Opportunity Cost}.

In \textbf{FS geometry}, opportunity cost is defined by \textbf{orthogonal subspaces}.

When you decide to ``accompany your loved one'' in this second, what do you do geometrically?

You project your wave function onto the basis of $|Love\rangle$.

At the same time, you must forcibly \textbf{zero out} the projection coefficients of all other bases orthogonal to $|Love\rangle$ (such as $|Work\rangle$, $|Travel\rangle$, $|Sleep\rangle$).

\begin{itemize}
\item \textbf{You have killed countless ``possible selves.''}

\item Those versions of you in parallel universes who might become billionaires, who might become vagabonds---in this second, they are all \textbf{sacrificed} by you.
\end{itemize}

You use countless ``unhappened futures'' as fuel to mint this single ``happened present.''

This is why true love is expensive, why focus is difficult.

Because every choice is a \textbf{massacre}.

\subsection{The Anchor of Value: Irreversibility}

If we could save and load (Time Travel), scarcity would disappear.

If I choose A and find it unsatisfactory, I can load back and choose B, then both A and B become cheap.

But \textbf{$\pi$ (inertia/conservation)} locks this path.

In Volume One, we established the concept of \textbf{``crystallization''}. Once a transaction is completed, $v_{ext}$ transforms into $v_{int}$, the wave function collapses, and history is \textbf{``written dead''} on the four-dimensional manifold.

\textbf{Irreversibility} is the ultimate anchor of value.

\begin{itemize}
\item This second passes, and it will never return.

\item These words are spoken, and they can never be taken back.

\item This person is missed, and you will never meet them again.
\end{itemize}

It is precisely this \textbf{``one-time''} property that endows time with divinity.

The universe is a \textbf{``Hardcore Mode''} game, with only one life, no save files.

Therefore, every action is \textbf{epic}.

\subsection{Conclusion: You Are Yourself a Luxury}

When we understand scarcity, we will no longer feel bored by ``ordinary daily life.''

The ten minutes you spend daydreaming by the window are not geometrically blank.

They are the \textbf{unique manifestation} carefully selected by the universe from infinite-dimensional possibilities.

They are ten minutes of tranquility you won for yourself by consuming enormous $c_{FS}$ budget in a war against entropy increase.

\textbf{You are the most luxurious existence in the universe.}

Because to let you experience this second of ``boredom,'' the entire universe is frantically burning computational power in the background, calculating countless parallel worlds you abandoned, just to support this single certain reality before your eyes.

Since we know that the value of time comes from scarcity, then in this market where an inch of gold cannot buy an inch of time, what kind of ``goods'' are worth exchanging? What is the only asset in the universe that \textbf{does not depreciate}?

This leads to the theme of the next section: \textbf{Negentropy as Hard Currency}. We will see that among all values, only one thing can traverse time's inflation, and that is \textbf{order}.

