\chapter{Open Universe Dynamics --- Mathematical Models of Capacity Growth and Red Queen Games}

This appendix provides dynamical proofs for Chapter 9 ``Rejecting Heat Death'' and Chapter 12 ``No Omega Point.'' We will construct coupled equations describing the evolution of cosmic capacity, entropy, and agent complexity over time.

\section{Dynamic Bekenstein Bound}

Consider an expanding universe dominated by dark energy $\Lambda$.

Although the event horizon is fixed, for computational cosmology, what matters more is the \textbf{number of computational nodes (QCA lattice points) within comoving volume} $N(t)$.

In QCA expansion model (node addition model):

$$N(t) \propto a(t)^3 \propto e^{3Ht}$$

where $H = \sqrt{\Lambda/3}$ is the Hubble constant.

The universe's potential \textbf{information capacity upper limit} $I_{max}(t)$ grows with $N(t)$:

$$I_{max}(t) \sim N(t) \sim e^{3Ht}$$

\section{Limits on Entropy Production}

The universe's \textbf{used information (entropy)} $S(t)$ mainly comes from matter particle decoherence and black hole radiation.

Since total matter is conserved (or grows slower than volume), and computation processes are limited by light speed (Bremermann Limit):

$$\frac{dS}{dt} \le \frac{E_{total}}{\hbar} \cdot N_{particles}$$

In a sparse universe, $S(t)$ growth is at most linear or polynomial, far below $I_{max}(t)$'s exponential growth.

\textbf{Corollary B.1 (Infinite Blank Theorem)}:

$$\lim_{t \to \infty} \frac{S(t)}{I_{max}(t)} = 0$$

The universe's \textbf{Filling Ratio} tends to zero. This means the universe is forever ``new,'' and heat death ($S \to I_{max}$) is mathematically impossible.

\section{Red Queen Game Equations}

We use a variant of \textbf{Lotka-Volterra equations} to describe the game between agents ($A$) and environmental entropy ($E$).

\begin{itemize}
\item \textbf{Agent complexity ($C$)}: Represents $M_I$.

\item \textbf{Environmental entropy pressure ($P$)}: Represents survival difficulty.
\end{itemize}

$$
\begin{cases}
\frac{dC}{dt} = \alpha C \cdot P - \beta C^2 & (\text{pressure drives evolution, but limited by metabolic cost}) \\
\frac{dP}{dt} = \gamma C - \delta P & (\text{intelligent activity produces waste heat increasing pressure, expansion dilutes pressure})
\end{cases}
$$

\begin{itemize}
\item \textbf{$\alpha C P$}: Red Queen term. The harsher the environment ($P$ large), the faster agents evolve to survive.

\item \textbf{$\gamma C$}: Landauer waste heat term. Higher intelligence emits more entropy, pushing $P$ higher.

\item \textbf{$-\delta P$}: Cosmic expansion term. Rapid space expansion dilutes waste heat.
\end{itemize}

\textbf{Phase Plane Analysis}:

When $\delta$ (expansion rate) is sufficiently large, this system has no stable fixed point (death), but forms \textbf{Limit Cycles} or \textbf{eternally diverging spirals}.

\textbf{Conclusion}:

As long as cosmic expansion ($\delta > 0$) can timely remove waste heat, agents can achieve \textbf{infinite exponential growth} in complexity through continuous arms races ($\alpha, \gamma$ cycles).

\textbf{Civilization's immortality is built on the foundation of cosmic expansion.}

