\section{Landauer's Principle and Thermodynamic Cost of Irreversible Computation}

In QCA discrete ontology, microscopic dynamics $U$ is strictly unitary (reversible). This means information never loses at microscopic level. However, computation in macroscopic world is often \textbf{logically irreversible}. Most typical example is "erase" operation (RESET): regardless of input being 0 or 1, output is forced to 0.

There is profound conflict between logical information loss (many-to-one mapping) and physical microscopic reversibility (one-to-one mapping). Rolf Landauer resolved this conflict in 1961, pointing out: \textbf{Logical irreversibility must be accompanied by physical heat dissipation}. This section will strictly derive Landauer principle from phase space volume conservation of QCA, and establish ontological status of "information as physics."

\subsection{Logical Irreversibility and Phase Space Compression}

Consider a physical bit (such as spin or QCA lattice point), its logical state space is $\mathcal{S} = \{0, 1\}$.

\begin{itemize}
\item \textbf{Logically Reversible Operation} (such as NOT): $0 \to 1, 1 \to 0$. This is bijection, entropy unchanged.

\item \textbf{Logically Irreversible Operation} (such as ERASE): $0 \to 0, 1 \to 0$. This is compression mapping, entropy decreases.
\end{itemize}

In QCA framework, microscopic states are described by state vectors $|\psi\rangle$ in Hilbert space.

\textbf{Quantum Version of Liouville's Theorem} states: unitary evolution $U$ preserves \textbf{volume} (or von Neumann entropy) of state space.
$$S(\rho_{final}) = S(U \rho_{initial} U^\dagger) = S(\rho_{initial})$$

If we want to perform logical erasure (transforming unknown initial state $\rho_{mix} = \frac{1}{2}(|0\rangle\langle 0| + |1\rangle\langle 1|)$ to definite ground state $|0\rangle\langle 0|$), system entropy must change from $S = k_B \ln 2$ to $S = 0$.

Due to total entropy conservation, this reduced $k_B \ln 2$ entropy must be transferred to \textbf{Non-computational Degrees of Freedom} outside the system, i.e., environmental heat bath.

\subsection{Strict Derivation of Landauer Bound}

\begin{theorem}[Landauer's Principle]
\label{thm:landauer}
In any thermal environment at temperature $T$, minimum energy cost $\Delta E$ required to erase 1 bit of information (i.e., resetting physical system with Shannon entropy 1 bit to pure state) is:
$$\Delta E \ge k_B T \ln 2$$

Or expressed as minimum heat produced $Q$:
$$Q \ge k_B T \ln 2$$
\end{theorem}

\textbf{Proof}:

Let total system be $\text{System} \otimes \text{Bath}$.

Initial state: System in maximum mixed state $\rho_S = \mathbb{I}/2$, heat bath in thermal equilibrium $\rho_B = e^{-\beta H_B} / Z$.

Total entropy: $S_{tot} = S(\rho_S) + S(\rho_B) = k_B \ln 2 + S_B(E_B)$.

Process: Execute unitary operation $U_{SB}$ to achieve erasure.

Final state: System in pure state $\rho'_S = |0\rangle\langle 0|$, heat bath in new state $\rho'_B$.

Due to unitarity, total entropy conserved:
$$S(\rho'_S) + S(\rho'_B) \ge S_{tot} \quad (\text{equality if no additional correlations})$$

Substituting $S(\rho'_S) = 0$, we get:
$$S(\rho'_B) \ge k_B \ln 2 + S_B(E_B)$$

Entropy increase of heat bath $\Delta S_B \ge k_B \ln 2$.

According to thermodynamic definition $dS = dQ/T$, heat absorbed by heat bath (i.e., energy dissipated by system) is:
$$Q = \int T dS \ge T \Delta S_B \ge k_B T \ln 2$$

$\square$

\subsection{Garbage Bits and Heat Dissipation in QCA}

In QCA discrete networks, there is no abstract "heat bath." Landauer principle manifests as propagation of \textbf{Garbage Information}.

\begin{construction}[Reversible Embedding]
\label{constr:reversible}
Any irreversible logic gate (such as AND gate, $(a, b) \to a \land b$) can be embedded into a reversible gate (such as Toffoli gate) by adding auxiliary bits (Ancilla).
$$\text{Toffoli}(a, b, c) = (a, b, c \oplus (a \land b))$$

If we initialize $c$ to $0$, third bit of output is computation result. However, first two bits $(a, b)$ still remain at output, becoming \textbf{garbage bits}.

To perform next computation (reset register), we must "remove" these garbage bits.

In QCA, this manifests as \textbf{radiating} wave packets carrying garbage information to infinity (or into horizon). These radiated wave packets carry energy $E = \hbar \omega$.

For wave packet with frequency $\omega$, entropy it carries is $S$. To carry away 1 bit entropy, wave packet must have sufficient phase space volume. At thermal equilibrium, this corresponds to energy $E \ge k_B T \ln 2$.
\end{construction}

\textbf{Physical Corollary}:

CPU heating is not due to electron friction (that's engineering defect), but due to \textbf{reset of logic gates}. Each forced reset of $0 \to 0$ or $1 \to 0$ microscopically emits a photon (or phonon) into environment, carrying away entropy originally stored in bit.

\subsection{Physical Reality of Information: Exorcism of Maxwell's Demon}

Landauer principle completely resolves \textbf{Maxwell's Demon} paradox that plagued physics for a century.

Demon attempts to control gate by acquiring information about molecules (velocity), thereby reducing system entropy without doing work.

Landauer pointed out: Demon's brain (or memory) is a physical system. To work continuously, demon must constantly \textbf{erase} old memories to store new information.

\begin{itemize}
\item \textbf{Measurement Phase}: Entropy transfers from gas to demon's memory (gas entropy decreases, demon entropy increases).

\item \textbf{Erasure Phase}: Demon resets memory, according to Landauer principle, must release heat $Q \ge k_B T \ln 2$ to environment.
\end{itemize}

Entropy increase $\Delta S_{env} \ge k_B \ln 2$ produced by this heat precisely cancels entropy reduced by demon through sorting molecules.

\begin{conclusion}[Information as Physical Resource]
\label{concl:information-physical}
Information is not an abstract concept independent of matter; it is a form of \textbf{Negentropy}.

\begin{enumerate}
\item \textbf{Energy Cost}: Processing information (especially erasure) must consume free energy.

\item \textbf{Mass}: If $E=mc^2$, then erasing 1 bit information corresponds to increasing environmental mass $\Delta m = \frac{k_B T \ln 2}{c^2}$.

\item \textbf{Gravity}: Aggregation of large amounts of information (high entropy states) produces gravitational effects (black holes).
\end{enumerate}

This principle establishes thermodynamic bottom line of computational physics. In the next section 25.2, we will explore how nature utilizes \textbf{Fibonacci Coding} and \textbf{Zeckendorf Arithmetic} to achieve optimal local information transmission without erasing information (reversible computation).
\end{conclusion}

