\section{Appendix C: Selected Bibliography}

\begin{quote}
\textbf{Note}: Although this book is a theoretical work, its core arguments are built upon solid foundations of cutting-edge research in physics, information theory, and cognitive science. The following literature provides original mathematical proofs and experimental evidence for key concepts covered in the book (such as QCA, holographic principle, IIT, free energy principle). For readers wishing to explore technical details in depth, these works are essential paths into the depths of the ``rabbit hole.''
\end{quote}

\subsection{1. Physics Foundations \& Cosmology}

\begin{itemize}
\item \textbf{'t Hooft, G. (2016).} \textit{The Cellular Automaton Interpretation of Quantum Mechanics}. Springer.

\begin{itemize}
\item \textbf{Relevant Sections}: Foreword, 1.1, 3.3

\item \textbf{Core Contribution}: Foundational work by Nobel laureate Gerard 't Hooft, arguing that quantum mechanics' probabilistic nature may originate from underlying deterministic cellular automaton (QCA) evolution, providing the physics foundation for this book's ``discrete ontology.''
\end{itemize}

\item \textbf{Susskind, L. (1995).} ``The World as a Hologram''. \textit{Journal of Mathematical Physics}, 36(11), 6377--6396.

\begin{itemize}
\item \textbf{Relevant Sections}: 2.2, 3.2

\item \textbf{Core Contribution}: Foundational work on the holographic principle. Argues that physical information in three-dimensional space is completely encoded on two-dimensional boundaries, providing mathematical support for the view that ``physical world is a projection of information.''
\end{itemize}

\item \textbf{Maldacena, J., \& Susskind, L. (2013).} ``Cool horizons for entangled black holes''. \textit{Fortschritte der Physik}, 61(9), 781--811.

\begin{itemize}
\item \textbf{Relevant Sections}: 3.2, A.2.2

\item \textbf{Core Contribution}: Proposed the famous \textbf{ER=EPR conjecture}, proving the geometric equivalence of quantum entanglement (EPR) and spacetime wormholes (ER), which is the core basis for this book's view of ``love/entanglement'' as the source of gravity.
\end{itemize}

\item \textbf{Bekenstein, J. D. (1981).} ``Universal upper bound on the entropy-to-energy ratio for bounded systems''. \textit{Physical Review D}, 23(2), 287.

\begin{itemize}
\item \textbf{Relevant Sections}: 10.2, A.3.1

\item \textbf{Core Contribution}: Bekenstein Bound. Proved the upper limit of information storage in finite space, deriving the physicality of information.
\end{itemize}
\end{itemize}

\subsection{2. Consciousness, Information \& Complexity}

\begin{itemize}
\item \textbf{Tononi, G. (2004).} ``An information integration theory of consciousness''. \textit{BMC Neuroscience}, 5(1), 42.

\begin{itemize}
\item \textbf{Relevant Sections}: 6.1, A.1.1

\item \textbf{Core Contribution}: Original paper on Integrated Information Theory (IIT). Defined $\Phi$ value as a measure of consciousness level, supporting this book's panpsychist position that ``consciousness is an intrinsic property of systems.''
\end{itemize}

\item \textbf{Friston, K. (2010).} ``The free-energy principle: a unified brain theory?''. \textit{Nature Reviews Neuroscience}, 11(2), 127--138.

\begin{itemize}
\item \textbf{Relevant Sections}: 4.1, A.1.1

\item \textbf{Core Contribution}: Proposed the free energy principle. Explained how living organisms resist entropy increase by minimizing prediction error (variational free energy), which is the theoretical source for this book's view of ``pain as error signal.''
\end{itemize}

\item \textbf{Bennett, C. H. (1982).} ``The thermodynamics of computation---a review''. \textit{International Journal of Theoretical Physics}, 21(12), 905--940.

\begin{itemize}
\item \textbf{Relevant Sections}: 7.2, 10.2

\item \textbf{Core Contribution}: Connected Landauer's Principle with Maxwell's Demon, clarifying the equivalence between information erasure and thermodynamic work, which is the physical foundation for this book's ``anti-entropy technology.''
\end{itemize}
\end{itemize}

\subsection{3. Computation, Logic \& Mathematics}

\begin{itemize}
\item \textbf{Turing, A. M. (1936).} ``On Computable Numbers, with an Application to the Entscheidungsproblem''. \textit{Proceedings of the London Mathematical Society}, 2(42), 230--265.

\begin{itemize}
\item \textbf{Relevant Sections}: 11.2

\item \textbf{Core Contribution}: Definition of Turing machines and undecidability of the halting problem. Supports this book's assertion that ``cosmic algorithm never halts.''
\end{itemize}

\item \textbf{Gödel, K. (1931).} ``Über formal unentscheidbare Sätze der Principia Mathematica und verwandter Systeme I''. \textit{Monatshefte für Mathematik und Physik}, 38, 173--198.

\begin{itemize}
\item \textbf{Relevant Sections}: 11.2

\item \textbf{Core Contribution}: Incompleteness theorems. Proved that truth always exceeds provability, providing logical guarantee for ``open-ended finale'' and ``infinite truth.''
\end{itemize}

\item \textbf{Lloyd, S. (2000).} ``Ultimate physical limits to computation''. \textit{Nature}, 406(6799), 1047--1054.

\begin{itemize}
\item \textbf{Relevant Sections}: 3.3, 10.2

\item \textbf{Core Contribution}: Calculated the maximum computational power limit of the universe as a computer, quantifying physical evolution as logical operations.
\end{itemize}
\end{itemize}

\subsection{4. Philosophy \& Theology}

\begin{itemize}
\item \textbf{Teilhard de Chardin, P. (1955).} \textit{The Phenomenon of Man}. Harper \& Row.

\begin{itemize}
\item \textbf{Relevant Sections}: 6.2, 10.1

\item \textbf{Core Contribution}: Proposed the concept of \textbf{Omega Point}, that the universe is evolving toward an extremely complex super-consciousness state, which is the core inspiration for this book's view on the ultimate destination of civilizations.
\end{itemize}

\item \textbf{Spinoza, B. (1677).} \textit{Ethics}.

\begin{itemize}
\item \textbf{Relevant Sections}: 0.2, 9.1

\item \textbf{Core Contribution}: Spinoza's pantheism (Deus sive Natura), that God equals nature/universe as a whole.
\end{itemize}

\item \textbf{Spencer-Brown, G. (1969).} \textit{Laws of Form}. Allen \& Unwin.

\begin{itemize}
\item \textbf{Relevant Sections}: 2.1

\item \textbf{Core Contribution}: Proposed ``distinction'' as the starting point of cognition, which is the logical prototype for this book's view of ``Big Bang as dissociation.''
\end{itemize}
\end{itemize}

---

\textbf{(End of Book)}

