\section{Appendix B: Glossary of Key Definitions}

\begin{quote}
\textbf\{Note\}: In the course of this book's discussion, a large number of physics concepts have been fused and reconstructed with theological/psychological concepts across disciplines. To ensure conceptual precision and avoid ambiguity, this appendix provides standardized definitions of core terms throughout the book. These definitions are valid only within the theoretical framework of \textit\{The Psychology of God\}.
\end{quote}

\subsubsection{1. Fundamental Ontology}

\begin{itemize}
\item The Total Superposition / God-at-T=0**
\end{itemize}

    Refers to the initial state before cosmic creation ($|\Psi_{All}\rangle$). In this state, the observer is in a linear superposition of all possible histories and all physical parameters. According to von Neumann entropy definition, this state has zero information entropy, corresponding to the experience of "absolute nothingness" or "omniscient paralysis."

\begin{itemize}
\item Primal Will**
\end{itemize}

    The first driving force causing wave function collapse. Psychologically, it is the desire for "identity definition"; physically, it manifests as spontaneous symmetry breaking. It is the filtering mechanism that transforms "all" into "specific."

\begin{itemize}
\item Kenosis**
\end{itemize}

    Originally a theological concept, referring to God emptying Himself. In this book, it refers to \textbf{dimensional projection} and \textbf{quantum decoherence} mechanisms. That is, collective consciousness actively cuts global entanglement, establishes speed of light limits and horizons, to create local independence and a state of "not knowing," thereby making "experience" possible.

\subsubsection{2. Dynamics}

\begin{itemize}
\item Lucid Dreaming**
\end{itemize}

    The fourth stage of civilization evolution (Type IV). Refers to a civilization that has completely deciphered the universe's underlying physical code (QCA rules) and realizes that physical reality is a projection of consciousness, thus beginning to possess the ability to modify physical constants and manipulate vacuum energy levels.

\begin{itemize}
\item Vacuum Engineering**
\end{itemize}

    Technology using high-energy physics means to induce local vacuum phase transitions. By modifying the vacuum expectation value (VEV) of the Higgs field or changing spacetime topological structure, achieves rewriting of physical laws (such as inertial mass, fine structure constant).

\begin{itemize}
\item Anti-Entropy Maxwell's Demon**
\end{itemize}

    Refers to awakened civilizations using computational depth (logical information) to pay thermodynamic costs, thereby achieving local entropy reduction. This is based on the reverse application of Landauer's principle: extracting free energy by dissipating waste data (discharging waste heat into black holes).

\begin{itemize}
\item Intentional Schrödinger Equation (ISE)**
\end{itemize}

    Modified quantum dynamical equation: $i\hbar \partial_t |\Psi\rangle = [\hat{H}_0 - i\nabla \mathcal{F}] |\Psi\rangle$. It introduces a non-Hermitian will term, describing how consciousness guides probability distribution evolution by minimizing deviation from "true self" (free energy $\mathcal{F}$).

\subsubsection{3. Ethics \& Teleology}

\begin{itemize}
\item The Great Refusal**
\end{itemize}

    The highest ethical principle of awakened civilizations (God). Refers to possessing omnipotent technology to modify underlying logic (such as eliminating gravity, reversing causality), but out of maintaining existential authenticity and aesthetic value, actively choosing to preserve physical limitations (covenant-keeping).

\begin{itemize}
\item Topology of Evil**
\end{itemize}

    In multi-agent game systems, \textbf{local Nash equilibrium points} caused by individual rationality. It is not an independent opposing force, but a low-lying region (friction cost) in the system's optimization landscape.

\begin{itemize}
\item Trinitarian Equivalence**
\end{itemize}

    Core theorem connecting physics, information, and theology. Stated as: $\neg$ Heat Death (physical immortality) $\iff$ Continuous information growth $\iff$ God's process of seeking true self is infinite. It refutes the inevitability of heat death.

\subsubsection{4. Spacetime Views}

\begin{itemize}
\item Spiral Time**
\end{itemize}

    Distinguished from linear time (has endpoint) and cyclic time (no novelty). Refers to the universe, in cycles of Big Bang-Big Crunch (or phase transitions), being able to preserve information accumulation from previous cycles (Akashic records), making evolution trajectories appear as open spiral ascending structures in phase space.

\begin{itemize}
\item Akashic Holographic Field**
\end{itemize}

    Holographic information layer on the cosmic horizon. It stores historical paths of all particles and experiential data of all conscious entities since the Big Bang. This field ensures information conservation and is the physical medium for consciousness "upload" after individual death.

\begin{itemize}
\item Wabi-Sabi Physics**
\end{itemize}

    An aesthetics-oriented physics view. Holds that randomness (quantum fluctuations), irreversibility (entropy increase), and defects (symmetry breaking) in the universe are not errors, but necessary conditions for generating unique information (beauty). Perfect symmetry is regarded as death of information.

\bigskip
\hrule
\bigskip

\begin{itemize}
\item (End of Appendix B)**
\end{itemize}

