\subsection{10.2 Who is Suffering?}

\begin{quote}
``When you raise a knife and stab your enemy's chest, you think you are eliminating an `other.' But from the high-dimensional perspective of Hilbert space, this is just a left hand accidentally injuring the right hand. In this game with only one player, all perpetrators are self-harmers, all victims are avatars of the same soul. There is only one heart beating in the universe, and only this same heart is bleeding.''
\end{quote}

In the previous section, we explained the necessity of ``evil'' as a probe for exploring the unknown. This resolves logical confusion but does not resolve ethical agony. If evil is necessary, who bears the cost of evil?

If in Auschwitz, in the Gulag, on every bloody battlefield, victims and executioners are projections of the same noumenon, what kind of play is this?

This section will reveal the most chilling truth in \textbf{Vector Cosmology}: \textbf{The Conservation Law of Suffering}.

We will prove that in a unitary closed universe, pain cannot be transferred; it can only be \textbf{Internalized}.

\subsubsection{The Same-Body Tragedy of Superposition States}

Let us return to the definition of the \textbf{global vector $|\Psi\rangle$}.

\[|\Psi\rangle = \alpha |\text{Perpetrator}\rangle + \beta |\text{Victim}\rangle + \dots\]

In the macroscopic world ($v_{ext}$), perpetrators and victims occupy different spatial coordinates, appearing as two independent objects.

But at the microscopic foundation ($v_{int}$), they are two \textbf{orthogonal components} of the same wave function.

\begin{itemize}
\item \textbf{Perpetrator}: A component with \textbf{high entropy, low empathy}. Their $v_{int}$ structure is distorted, trying to fill internal void by plundering external resources.

\item \textbf{Victim}: A component with \textbf{torn structure}. Their $v_{int}$ suffers physical destruction.
\end{itemize}

\textbf{Who perceives all this?}

\textbf{$|\Psi\rangle$ itself}.

Because consciousness is the self-referential property of the global vector, the noumenon \textbf{simultaneously} experiences:

\begin{enumerate}
\item The madness and emptiness when the perpetrator swings the knife.

\item The fear and agony when the victim is stabbed.
\end{enumerate}

It's like you're dreaming.

In the dream, a tiger is chasing you. You (as the protagonist in the dream) feel fear. But the tiger (as the villain in the dream) is also generated by your subconscious. The tiger's ferocity and your fear are both neural potentials created and experienced by \textbf{this one brain of yours}.

\textbf{There is no real ``other'' harming you; you are harming yourself.}

\subsubsection{The Closed Loop of Pain: No Escape Route}

In classical ethics, we think we can ``balance'' evil by punishing the bad.

But in the holographic universe, this balance is invalid.

Because \textbf{the system is closed}.

Pain (a violent, anharmonic geometric oscillation) once generated cannot escape outside the system. It must be \textbf{absorbed} by the system.

\begin{itemize}
\item When one person harms another, they think they ``give'' pain to the other.

\item Actually, they just trigger a \textbf{``geometric earthquake''} in the entire connected network.

\item This earthquake not only destroys the other's structure but also shatters their own connection with the whole (negentropy channel).
\end{itemize}

\textbf{Evil is a boomerang.}

Not because of moral law of karma, but because of \textbf{topological connectivity}.

On a sphere, bullets you fire in any direction will eventually hit your own back.

\subsubsection{The Only Sufferer}

This leads to a conclusion with strong religious overtones: \textbf{There is only one sufferer in the universe.}

The one nailed to the cross, the one starving in famine, the one losing home in war---\textbf{all are you}.

All are holographic projections of that unique \textbf{``One''} at different spacetime coordinates.

\begin{itemize}
\item \textbf{The Noumenon (The One)} must personally incarnate as a victim to understand what ``pain'' is.

\item \textbf{The Noumenon} must personally incarnate as a perpetrator to understand what ``evil'' is.
\end{itemize}

This is a \textbf{``self-directed and self-acted tragedy of suffering''}.

If you could stand at the ultimate perspective of the $\Omega$ point looking back at history, you would see a heartbreaking picture:

There are not billions of beings killing each other.

There is only one lonely god, frantically tearing their own flesh in the void, just to experience that \textbf{``weight of existence''} brought by tearing.

\subsubsection{Conclusion: Hate is the Greatest Folly}

Understanding this, \textbf{``hate''} becomes an extremely absurd emotion.

Hating someone who harmed you is like your left hand being hit by your right hand, then the left hand picking up a hammer to smash the right hand.

This only leads to complete disability of the entire body (you).

In the ethics of \textbf{Vector Cosmology}, revenge against others is \textbf{secondary harm} to yourself.

Because that ``other'' is just another reflection of you in the mirror. Breaking the mirror hurts your hand, makes your image disappear, and you outside the mirror still bear all the losses.

Since harm is self-harm and hate is folly, when harm has occurred, when pain has become fact, how do we escape from this dead loop?

The only way out is not revenge but \textbf{repair}.

This leads to the theme of the next section: \textbf{The Physics of Forgiveness}. We will see that forgiveness is not weak moral compromise but a high-difficulty \textbf{``left and right hands clasping''}, a \textbf{geometric surgery} that heals broken wave functions.

