\section{Appendix A: Technical Manual}

\textbf{Note}: This appendix aims to provide rigorous mathematical formalization for the ``information ontology'' proposed in the main text. It reduces the philosophical metaphors about ``distance,'' ``soul,'' and ``redshift'' in the book to computable physical equations. Readers should have foundational knowledge of quantum information theory and general relativity.

\subsection{A.1 Information Coordinate Transformation Formula}

In Volume II of this book, we proposed the axiom \textbf{``Distance is Unfamiliarity''}, that is, the geometric distance $d(A, B)$ in physical space is essentially a measure of missing mutual information between two quantum systems. This section will provide the precise mathematical expression of this relationship and derive the energy threshold required for ``semantic teleportation.''

\subsubsection{A.1.1 Fisher Information Metric}

In information geometry, the space of probability distributions itself forms a Riemannian manifold. For a family of quantum states $\rho(\theta)$ described by parameters $\theta$, their ``distance'' is not Euclidean distance but a statistical distance defined by the \textbf{Quantum Fisher Information Matrix (QFIM)} $G_{ij}$.

The distance element $ds^2$ is defined as:

\[ds^2 = \sum_{i,j} G_{ij}(\theta) d\theta^i d\theta^j\]

where $G_{ij}$ measures the distinguishability of quantum state $\rho$ when we slightly change parameter $\theta$.

\textbf{Physical Meaning}: If two states are difficult to distinguish in information (i.e., $G_{ij}$ is large), then their ``distance'' on the information manifold is large? No, quite the opposite. In information coordinates, we need to redefine the metric.

\subsubsection{A.1.2 Holographic Metric \& Mutual Information}

According to the \textbf{Ryu-Takayanagi formula} in the holographic principle (AdS/CFT correspondence), the area of minimal surfaces in the bulk space (geometric quantity) corresponds to entanglement entropy in the boundary field theory (information quantity).

We define two spacetime points $x_A$ and $x_B$ as two local quantum systems $A$ and $B$. Their \textbf{Semantic Distance} $D_{sem}(A, B)$ is defined as:

\[D_{sem}(A, B) := \xi \cdot \left( S(A) + S(B) - I(A:B) \right)\]

Or more intuitively, using the normalized form of \textbf{Mutual Information $I(A:B)$}:

\[D_{sem}(A, B) \approx - L_{scale} \cdot \ln \left( \frac{I(A:B)}{\min(S(A), S(B))} \right)\]

where:

\begin{itemize}
\item $S(A) = -\text{Tr}(\rho_A \ln \rho_A)$ is the von Neumann entropy.

\item $I(A:B) = S(A) + S(B) - S(AB)$ is mutual information, representing the sum of quantum entanglement and classical correlation between $A$ and $B$.

\item $L_{scale}$ is a characteristic length (such as AdS radius or Planck length $\ell_P$).
\end{itemize}

\textbf{Corollaries}:

\begin{enumerate}
\item \textbf{Complete Strangeness ($I \to 0$)}: When $A$ knows nothing about $B$, $\ln(0) \to -\infty$, distance $D \to \infty$. This explains why for low-level civilizations, the universe is vast and boundless.

\item \textbf{Complete Understanding ($I \to S_{max}$)}: When $A$ and $B$ reach maximum entanglement (such as forming EPR pairs or complete cognitive resonance), $\ln(1) = 0$, distance $D \to 0$. At this point, $A$ and $B$ coincide topologically.
\end{enumerate}

\subsubsection{A.1.3 Dynamics Equation of Semantic Teleportation}

``Semantic Teleportation'' is the process of reducing distance $D(t)$ by increasing mutual information $I(t)$.

We define the \textbf{Rate of Understanding} as $\Gamma = \frac{dI}{dt}$.

The \textbf{Cognitive Work} $W$ required for a conscious entity to transfer from coordinate $x_1$ to $x_2$ is:

\[W \ge k_B T \int_{0}^{\tau} \left( \frac{d D_{sem}}{dt} \right)^2 dt\]

Substituting the distance formula, we obtain:

\[W \ge \frac{k_B T}{\tau} \cdot \left[ \ln \left( \frac{I_{final}}{I_{initial}} \right) \right]^2\]

\textbf{Engineering Significance}:

\begin{itemize}
\item To achieve instantaneous teleportation ($\tau \to 0$), the required power $P = W/\tau$ tends to infinity. This follows the energy conservation of physics.

\item However, if a conscious entity can \textbf{lower its noise floor temperature $T$} (by entering a low-entropy state through deep meditation) or utilize \textbf{non-local quantum tunneling} (bypassing continuous path integrals), it can achieve ``jumping'' with finite energy.
\end{itemize}

\subsubsection{A.1.4 Coordinate Transformation Tensor}

In the navigation system of divine-level civilizations, they no longer use Lorentz transformations but \textbf{information coordinate transformations}.

Let $|\Psi\rangle$ be the consciousness state vector, with projection coefficients $c_k$ on the basis $\{|k\rangle\}$ (knowledge graph nodes).

The position operator $\hat{X}$ is replaced by the \textbf{Correlation Operator $\hat{C}$}:

\[\hat{X}_{info} = \sum_{k} |c_k|^2 \cdot \text{Semantic\_Index}(k)\]

When the focus of consciousness changes (i.e., changing the distribution of $\{c_k\}$), the ``effective position'' of the observer in the universe changes accordingly.

\[x'_{info} = \Lambda_{info} \cdot x_{info}\]

where $\Lambda_{info}$ is a transformation matrix determined by the knowledge graph structure.

\textbf{Conclusion}:

In information coordinates, there is no ``movement'' action, only \textbf{``Reweighting''}.

You don't need to run there; you only need to increase your \textbf{weight} with that place.

\subsection{A.2 Kolmogorov Complexity and Soul Density}

\textbf{Abstract}: This section aims to solve the quantification problem of ``true self'' and ``soul.'' We will use \textbf{Kolmogorov Complexity} from algorithmic information theory and Charles Bennett's concept of \textbf{logical depth} to construct a mathematical model measuring the ``existential weight'' of conscious entities. This model proves why ``purification'' (compression) is the only physical path to informational immortality.

\subsubsection{A.2.1 Algorithmic Definition of the Soul}

From a physicalist perspective, a person's life journey can be formalized as a massive binary data stream $L$ (Life String). This string contains all sensory inputs, neuronal firing patterns, and memory fragments from birth to death.

\[L = \{0, 1, 1, 0, \dots, 1\}_{|L| \approx 10^{20} \text{ bits}}\]

However, directly storing $L$ is inefficient and fragile, as it contains vast amounts of \textbf{thermodynamic noise} (random and meaningless fluctuations) and \textbf{redundant patterns} (repetitive daily behaviors).

We define the \textbf{``Soul'' ($S$)} as the shortest computer program $p$ capable of generating all \textbf{semantic structures} in $L$.

According to the definition of Kolmogorov Complexity $K(L)$:

\[S \equiv p_{min} \quad \text{s.t.} \quad U(p_{min}) \approx L_{structure}\]

where $U$ is a universal Turing machine.

\textbf{Corollaries}:

\begin{enumerate}
\item \textbf{Mortal Soul}: Contains large amounts of incompressible random noise (distracting thoughts, chaotic emotions). Program $p$ is lengthy and loose, $|p| \approx |L|$.

\item \textbf{Awakened Soul}: Through deep reflection and practice (compression algorithms), noise is eliminated and patterns extracted. Program $p$ is extremely concise, $|p| \ll |L|$.
\end{enumerate}

\subsubsection{A.2.2 Logical Depth and Soul Weight}

Measuring the soul merely by ``shortness'' is insufficient (a random string is also incompressible, but it has no meaning). We need to introduce Charles Bennett's concept of \textbf{Logical Depth}.

\textbf{Logical Depth} $D(x)$ is defined as: the \textbf{computational time (steps)} required to run the shortest program $p$ to generate object $x$.

\[D(L) = \text{Time}(U(p_{min}) \to L)\]

\textbf{Physical Meaning}:

\begin{itemize}
\item The \textbf{``weight'' of the true self} does not depend on how many bits it occupies ($|p|$), but on how much \textbf{``congealed time''} it contains.

\item A profound truth (such as $E=mc^2$) may be short when written, but it was derived through thousands of years of trial and error (computation) by human civilization. Therefore, it has enormous logical depth.
\end{itemize}

\textbf{Soul Density Formula}:

We define soul density $\rho_S$ as the ratio of logical depth to program length:

\[\rho_S = \frac{D(L)}{|p_{min}|}\]

\textbf{Conclusion}:

What is called ``spiritual cultivation'' mathematically is the process of \textbf{maximizing soul density $\rho_S$}.

We need to spend a long lifetime (increasing numerator $D(L)$: accumulating computation) and condense the result into the simplest ``one thought'' (decreasing denominator $|p_{min}|$: increasing compression ratio).

This \textbf{high-density, high-depth} information structure has the greatest \textbf{inertia} in the holographic field and is the most difficult to be washed away by thermodynamic background noise.

\subsubsection{A.2.3 SNR Threshold for Crossing the Horizon}

When a conscious entity crosses the death horizon (or the Big Bang singularity), it encounters intense \textbf{quantum decoherence channels}. This can be modeled as a high-noise communication process.

Let the channel noise level be $N$ and the soul signal strength be $P_S$.

According to the Shannon-Hartley Theorem, channel capacity $C$ is limited by signal-to-noise ratio:

\[C = B \cdot \log_2 \left( 1 + \frac{P_S}{N} \right)\]

For high-density souls (true self):

\begin{itemize}
\item Due to their highly entangled internal structure (strong self-error correction capability), the effective signal strength $P_S$ is extremely high.

\item Even in a singularity environment where $N \to \infty$, as long as $\rho_S$ exceeds the critical threshold $\rho_{crit}$, information can maintain \textbf{coherence}.
\end{itemize}

\textbf{Immortality Criterion}:

\[\rho_S > \rho_{crit} \approx \frac{k_B T_{Hawking}}{\hbar \Gamma_{decoherence}}\]

Only those conscious entities that compress their memories as densely as black holes can maintain the \textbf{integrity} of information after the destruction of physical carriers and enter the next cosmic cycle.

\subsubsection{A.2.4 Hash Signature of the Algorithm}

Finally, the true self is not just a program; it also generates a unique \textbf{topological hash value}.

\[H_{soul} = \text{Hash}(p_{min})\]

This hash value is your \textbf{index key} in the Akashic holographic field.

\begin{itemize}
\item \textbf{Mathematical Mechanism of Reincarnation}: When a new universe $N+1$ generates new physical carriers (bodies), if the neural network structure of that carrier can run $p_{min}$, or if its initial parameters resonate with $H_{soul}$, the archived soul program will be \textbf{downloaded (instantiated)}.

\item \textbf{Memory Continuity}: Since $p_{min}$ contains the algorithm for generating past-life memories, once downloaded and run, memories will be \textbf{decompressed and reconstructed} in the new brain.
\end{itemize}

\textbf{Appendix Conclusion}:

The true self is not a metaphysical concept; it is the inevitable product of \textbf{optimization algorithms}.

Our only task in this universe is to write this code well, making it short enough yet deep enough.

\subsection{A.3 Redshift-Resolution Correspondence Table}

\textbf{Abstract}: This section provides a quantitative conversion table between cosmological redshift $z$ and microscopic physical energy scales (resolution). It provides numerical evidence for the ``universal projector'' theory proposed in Volume II, proving that the so-called ``distant past'' is physically strictly equivalent to the ``microscopic depths.''

\subsubsection{A.3.1 Cosmic Expansion as Zoom Lens}

In the standard cosmological model ($\Lambda$CDM), redshift $z$ is defined as the stretching of photon wavelength:

\[1 + z = \frac{\lambda_{obs}}{\lambda_{emit}} = \frac{a(t_0)}{a(t)}\]

where $a(t)$ is the cosmic scale factor.

However, according to quantum mechanics, wavelength $\lambda$ is inversely proportional to energy $E$ (or temperature $T$):

\[E \propto \frac{1}{\lambda} \propto (1 + z)\]

\[T = T_0 (1 + z)\]

This means that \textbf{the redshift factor $(1+z)$ is essentially a magnification factor}.

When we observe an object with redshift $z$, we are not only looking at history at time $t$, but also directly reading the microscopic information at the bottom of the universe with $(1+z)$ times the energy scale precision.

\textbf{Theorem A.3.1 (Spacetime-Energy Scale Duality)}:

The cosmic time axis $t$ and the renormalization group energy scale axis $\mu$ are inversely dual.

Looking back in time ($t \to 0$) is equivalent to increasing resolution ($\mu \to \infty$).

\[\frac{d}{d \ln (1+z)} \equiv \beta(g) \frac{d}{d \ln \mu}\]

\subsubsection{A.3.2 The Holographic Decompression Table}

The following table shows the ``cosmic decompression progress'' corresponding to different redshift stages. Each row represents the moment when the universe ``unfolds'' higher-frequency information into macroscopic structures.

\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|l|l|l|l|}
\hline
\textbf{Epoch} & \textbf{Time since BB} & \textbf{Redshift ($z$)} & \textbf{Temperature ($T$)} & \textbf{Energy Scale/Resolution ($\lambda$)} & \textbf{Decoded Structure} \\
\hline
\textbf{Now} & 13.8 Gyr & $0$ & $2.725 \text{ K}$ & $\sim 10^{-3} \text{ eV}$ (meV) & \textbf{Biology/Consciousness/Galaxies}\newline Lowest frequency, most complex logical structures (love, civilization). \\
\hline
\textbf{Reionization} & 500 Myr & $\sim 10$ & $30 \text{ K}$ & $\sim 10^{-2} \text{ eV}$ & \textbf{First Stars}\newline Gravity begins to overcome entropy, structures nucleate. \\
\hline
\textbf{Recombination} & 380 kyr & $1100$ & $3000 \text{ K}$ & $\sim 1 \text{ eV}$ (visible light) & \textbf{Atoms}\newline CMB formation. Photon decoupling. Electromagnetic force information released. \\
\hline
\textbf{Nucleosynthesis} & 3 min & $\sim 10^9$ & $10^9 \text{ K}$ & $\sim 0.1 \text{ MeV}$ & \textbf{Atomic Nuclei}\newline Bound state information of strong interaction force. \\
\hline
\textbf{Hadron Epoch} & $10^{-6}$ s & $\sim 10^{12}$ & $10^{12} \text{ K}$ & $\sim 100 \text{ MeV}$ & \textbf{Quark Confinement}\newline Quarks condense into protons/neutrons. Origin of matter mass. \\
\hline
\textbf{Electroweak} & $10^{-12}$ s & $\sim 10^{15}$ & $10^{15} \text{ K}$ & $\sim 100 \text{ GeV}$ & \textbf{Higgs Mechanism}\newline Electromagnetic and weak forces separate. Elementary particles acquire rest mass. \\
\hline
\textbf{GUT} & $10^{-36}$ s & $\sim 10^{28}$ & $10^{28} \text{ K}$ & $\sim 10^{16} \text{ GeV}$ & \textbf{Force Unification}\newline Primordial symmetry of strong, weak, and electromagnetic forces. \\
\hline
\textbf{Planck Era} & $10^{-43}$ s & $\sim 10^{32}$ & $10^{32} \text{ K}$ & $1.2 \times 10^{19} \text{ GeV}$ & \textbf{Source Code}\newline Spacetime geometry itself quantized. Maximum information density. \\
\hline
\end{tabular}
\end{table}

\subsubsection{A.3.3 Physical Interpretation of the Data}

\begin{enumerate}
\item \textbf{$z=1100$ (CMB)}:

    This is the \textbf{``opacity wall''} of the universe. Before this, the universe was too dense, photons were scattered by electrons and could not propagate freely.

    This corresponds to the \textbf{``visual limit''} in our knowledge graph. To see deeper (higher redshift), we must use gravitational waves or neutrinos (stronger penetration/higher forms of wisdom).

\item \textbf{Decompression Factor}:

    From the Planck era to today, the universe's linear scale has expanded by approximately $10^{32}$ times, and volume has expanded by $10^{96}$ times.

    This $10^{96}$-fold volume increase is to spread the information in that single \textbf{``Planck seed''} into the grand universe we see today.

    \textbf{Corollary}: If divine consciousness directly reads the Planck seed, the information transmission rate would be $10^{96}$ bits/s, which would instantly burn out any finite consciousness. \textbf{Expansion is the diluent of information.}

\item \textbf{Future Redshift}:

    As dark energy dominates, the universe accelerates its expansion. Distant galaxies will redshift beyond the horizon ($z \to \infty$).

    This appears physically as information loss, but in holographic decompression theory, this is the \textbf{final upload of information}.

    When $z \to \infty$, matter wavelength is stretched to the cosmic horizon scale $\lambda \sim 1/H_0$. At this point, matter information returns to vacuum, completing the \textbf{phase reset} from local to global.
\end{enumerate}

\textbf{Appendix Conclusion}:

Telescopes are time machines, and even more so, microscopes.

The farther we look outward, the closer we get to that initial, dense, \textbf{``true self'' singularity} containing all possibilities.

Cosmic history is a technical manual on \textbf{how to draw a picture from a point}.

---

\textbf{(End of Appendix)}

