\subsection{9.1 Relativistic Ontology}

\begin{quote}
``In special relativity, there is no absolute rest; in vector cosmology, there is no absolute subject. In your coordinate system, you are the observer, I am the wave function; in my coordinate system, I am the observer, you are the wave function. This is like two spaceships flying away from each other at high speed; each sees the other's time slow down. This symmetry of `mutual objectivity' is the physical foundation of cosmic ethics.''
\end{quote}

\subsubsection{The Dictatorship of Reference Frames: Everyone is Their Own God}

Let us return to the basic operation of physics: establishing coordinate systems.

When you open your eyes to see the world, you are actually establishing an orthogonal coordinate system in Hilbert space with \textbf{``Self'' ($Self$)} as the origin.

\begin{itemize}
\item \textbf{Your position}: Always $(0, 0, 0)$. You are the center of the world.

\item \textbf{Your perspective}: You can only see things within your light cone.

\item \textbf{Your definition}: You measure the $v_{ext}$ (external attributes) of everything around you through your $v_{int}$ (internal value).
\end{itemize}

Within this reference frame, \textbf{Solipsism} is physically valid.

Because all data are input to \textbf{you}. All wave functions collapse for \textbf{you}. You have absolute interpretive power over this local universe.

\begin{itemize}
\item When you look at a person, they are just a collection of light and shadow (wave function) on your retina.

\item You cannot prove they have consciousness. You can only \textbf{``assume''} they have consciousness.

\item In strict geometric terms, for you, they are a \textbf{Projection} in your dream.
\end{itemize}

\textbf{Conclusion:}

In each person's subjective universe, they are the only \textbf{Creator}. They weave their reality with their own cognition.

\subsubsection{Mirror Symmetry: Mutual NPCs}

But this logic must also be \textbf{Symmetric}.

When you stand on a street corner looking at Person A, you think they are just a passing extra (NPC) in your grand life script.

But at this moment, Person A is also looking at you.

In their coordinate system, \textbf{they} are the protagonist. They are experiencing a thrilling adventure (maybe they just broke up or won the lottery). And you, for them, are just a blurry background texture on the street corner, a Person B not even worthy of a name.

This is \textbf{Relativistic Ontology}:

\begin{itemize}
\item \textbf{Frame A (you)}: $I_A = 1$ (subject), $I_B = 0$ (object/wave function).

\item \textbf{Frame B (them)}: $I_B = 1$ (subject), $I_A = 0$ (object/wave function).
\end{itemize}

These two facts are \textbf{simultaneously true}, just as ``your time slows down'' and ``my time slows down'' are simultaneously true in special relativity.

There is no ``God's perspective'' to judge who is the real protagonist.

\textbf{Everyone is a protagonist, and everyone is also a supporting character.} It depends on which basis you project onto.

\subsubsection{Collision: Tunnel Connection of Dreams}

Since everyone is dreaming, why are our dreams coherent? Why don't we pass through each other's bodies?

Because \textbf{Physical laws ($H$)} are the \textbf{communication protocol} shared by all dreams.

When two independent ``solipsistic universes'' meet in spacetime, a \textbf{``dream collision''} occurs.

In quantum mechanics, this is called \textbf{Interaction}.

\begin{itemize}
\item You extend your hand (your dream boundary expands).

\item They extend their hand (their dream boundary expands).

\item Two hands meet.
\end{itemize}

At this instant, two originally orthogonal Hilbert space subsystems undergo \textbf{``off-diagonal element coupling''}.

\[H_{int} = g \cdot (A^\dagger B + B^\dagger A)\]

Energy and information exchange at this contact point.

\begin{itemize}
\item \textbf{For you}: Their hand is warm. This proves they are not just light and shadow; they have a \textbf{thermodynamic core ($v_{int}$)} like you.

\item \textbf{For them}: Your hand is warm. This proves you are also a real dissipative structure, not a ghost in their hallucination.
\end{itemize}

\textbf{Conclusion:}

Reality is not a single-player game. Reality is a \textbf{P2P (Peer-to-Peer) network}.

There is no central server rendering the world. The world is a holographic image \textbf{``negotiated''} through high-frequency interactions by 7 billion (or more) independent ``graphics cards'' (consciousness).

We dream each other.

We mutually support each other's sense of reality.

Without others' observations, your wave function would also diffuse due to lack of boundary constraints.

\textbf{You are the scenery in others' dreams; this is the evidence of your existence.}

Since we have established the equality of ``mutual dreaming,'' how do these countless dreams weave together without becoming chaotic schizophrenia? Is there a larger net that encompasses all ``solipsistic universes''?

This leads to the theme of the next section: \textbf{Indra's Net}. We will see that individual dreams are just dewdrops on that infinitely vast cosmic web.

