\section{1.2 The Moment of Transaction}

\begin{quote}
``Time is not a continuously flowing river; time is a series of crisp sounds at the cash register. Every time you make a choice, every time you take action, you are making an irreversible transaction: you pay `possibility' and buy back `reality.' ''
\end{quote}

In the previous section, we established ``the rent of existence'': to maintain life structures, we must passively pay maintenance fees. But this is only the baseline of survival. As alchemists of \textbf{``Auric''}, we are more concerned with how to \textbf{appreciate}.

In the economic model of \textbf{Vector Cosmology}, the true value of time does not lie in how many seconds it has flowed (Chronos), but in how many \textbf{``effective transactions''} have occurred.

This section will reveal the subjective nature of time's passage: \textbf{the sense of time originates from transaction frequency.}

\subsection{Collapse is Payment: Converting Kinetic Energy into Assets}

In physics, what is ``something happened''?

From the perspective of quantum mechanics, an event equals \textbf{wave function collapse}.

From the perspective of FS geometry, an event equals \textbf{rotation of vector projection}.

When we face the future, what we have is \textbf{$v_{ext}$ (external velocity/degrees of freedom)}.

\begin{itemize}
\item This represents ``where I can go,'' ``what I can do.'' It is divergent, uncertain, the \textbf{cash flow} of possibility.
\end{itemize}

When we make a decision (such as deciding to read this book, deciding to love someone), we \textbf{lock} this cash flow.

\begin{itemize}
\item It transforms into \textbf{$v_{int}$ (internal structure/memory)}. It is convergent, definite, \textbf{minted gold coins}.
\end{itemize}

\[v_{ext} \xrightarrow{\text{Decision/Action}} v_{int}\]

This is \textbf{``the moment of transaction''}.

Every present moment is a \textbf{payment gateway}.

You stand at the counter, exchanging your precious $c_{FS}$ budget that could have gone anywhere into a specific, unchangeable history.

\begin{itemize}
\item You cannot go both left and right.

\item You must pay for the possibility of ``going right'' to buy the reality of ``going left.''
\end{itemize}

\subsection{The Relativity of Psychological Time: Transaction Density}

This mechanism perfectly explains why our perception of time differs drastically in different states.

\subsubsection{1. Flow and High-Frequency Transactions}

When you are fully focused on work, creation, or deep conversation with a loved one, you feel ``time flies.''

Why? Because you are conducting \textbf{high-frequency transactions}.

Your consciousness is making countless tiny collapses every second: judgment, reaction, correction, confirmation.

\begin{itemize}
\item \textbf{Transaction density is extremely high}.

\item Your $c_{FS}$ budget is frantically transforming from $v_{ext}$ to $v_{int}$.

\item You have \textbf{minted} a large number of gold coins (high-quality memories) in a short time.
\end{itemize}

Although physical time (Chronos) has passed for a long time, because you have been in the exciting cycle of ``payment-acquisition,'' your subjective feeling is coherent and compact.

\subsubsection{2. Boredom and Transaction Stagnation}

When you have nothing to do, daydream, or are forced to do repetitive labor, you feel ``time drags on.''

Why? Because \textbf{transactions have stopped}.

\begin{itemize}
\item No new information enters your $v_{int}$.

\item Your $v_{ext}$ has not transformed into any meaningful structure.

\item You are only paying ``the rent of existence'' (dissipation) without buying any goods.
\end{itemize}

\textbf{Boredom is the stagflation of the soul.}

You watch your $c_{FS}$ flowing away but cannot spend it. This geometric anxiety of ``having money but nowhere to spend'' is translated by the brain into ``lengthy'' and ``painful.''

\subsection{Rejecting Invalid Transactions}

At this payment gateway, the most tragic thing is not having no money (short lifespan), but \textbf{spending money recklessly}.

Many people's lives are full of \textbf{``invalid transactions''}.

\begin{itemize}
\item Scrolling through short videos: This is a \textbf{``false transaction''}. Although the screen flashes and information flows, it does not truly increase your $v_{int}$ complexity. You pay attention (the most expensive currency) and only buy back a pile of $v_{env}$ (waste) forgotten in the next second.

\item Indecision: This is \textbf{``transaction blockage''}. You stand at the counter, holding money, hesitating to place an order. As a result, time passes, money is eaten by inflation (aging), and you end up empty-handed.
\end{itemize}

\textbf{The Alchemist's Code of Action:}

\textbf{Place orders decisively.}

Don't be afraid of buying wrong (making mistakes is also an increase in $v_{int}$, called experience). The most terrible thing is \textbf{not buying}.

As long as transactions occur, time is minted into entities.

\subsection{Conclusion: You Are Your Shopping Cart}

Finally, when we reach the end of life and look back to settle accounts, who are we?

\textbf{We are the sum of all transaction records in this life.}

\begin{itemize}
\item Your body is what you traded with food.

\item Your wisdom is what you traded with diligence.

\item Your love is what you traded with vulnerability and trust.
\end{itemize}

Time itself is neutral; it is just a blank check.

It is through your repeated \textbf{``confirm payment''} at the payment gateway that you cash this check into a unique life.

Since we have understood how to accumulate value through transactions, where are these ``gold coins'' (memories and history) we bought back stored? Have they really dissipated with entropy increase as physicists say?

This leads to the theme of the next chapter: \textbf{The Anchoring of Value}. We will search for the most solid \textbf{``geometric safe''} that can resist the universe's mechanism of forgetting.

