\subsection{5.2 The Closed Loop of Existence}

\begin{quote}
``We have been desperately searching for the first mover of the universe, trying to find the finger that pressed the start button. But when we traverse the mist of spacetime and finally grasp that hand, we are surprised to find: it is our own hand. The universe is not a straight line shot from the past to the future; it is a circle that gives birth to itself. You are both effect and cause.''
\end{quote}

In the previous section, the Novikov self-consistency principle tells us: you cannot change the past because the past has already happened. This sounds like a limitation. But if we reverse this logic, it becomes a \textbf{``privilege of creation''}.

If the past must be self-consistent with the future, then \textbf{future states can ``reverse anchor'' past history}.

This section will explore the most puzzling and fascinating structure in time travel theory---\textbf{Bootstrap}, also known as \textbf{Ontological Paradox}.

We will reveal that the Big Bang is not a random physical event; it is an \textbf{``reverse instruction''} sent by future observers to the beginning of time to establish their own existence.

\subsubsection{The Bootstrap Paradox: Information Without Origin}

Imagine you go back in time and give a copy of \textit{Hamlet} to young Shakespeare. Shakespeare copies the book, publishes it, and it becomes a masterpiece. Hundreds of years later, you buy this book and take it back to the past.

\textbf{Question: Who actually wrote \textit{Hamlet}?}

Shakespeare got it from your hand, you bought it from Shakespeare. The content of this book seems to have \textbf{``no origin''}. It exists out of thin air in this closed-loop timeline.

This is \textbf{Bootstrapping}.

In \textbf{Vector Cosmology}, \textbf{the universe itself is this \textit{Hamlet}.}

\begin{itemize}
\item \textbf{Physical laws ($H$)}: So exquisite, as if customized for life. Who designed them?

\item \textbf{Initial parameters ($c_0$)}: So precise, just allowing stars to ignite. Who tuned them?
\end{itemize}

Traditional physics says this is ``coincidence'' or ``survivor bias of the multiverse.''

But closed-loop causality says: \textbf{No, this is what ``you'' brought back.}

Awakened observers (future civilizations), through deep understanding of physical reality, inject the information \textbf{``life must exist''} into the singularity moment at $t=0$ through retro-causal channels (delayed choice/phase interference).

This information becomes the \textbf{boundary conditions} of the Big Bang.

\subsubsection{Observation as Creation: Reverse Collapse}

This gives the \textbf{``Anthropic Principle''} an active physical mechanism.

Usually we think: because the universe is like this, we are here.

But in the closed loop: \textbf{Because we are here, the universe must be like this.}

Let us return to the evolutionary picture of \textbf{FS geometry}.

\begin{itemize}
\item \textbf{At $t=0$}: The cosmic wave function $|\Psi(0)\rangle$ is in an extremely chaotic superposition state. It contains countless possible physical laws, countless dead universes, and very few universes that allow life.

\item \textbf{At $t=now$}: You (the observer) open your eyes. You perform an \textbf{``existence measurement''}.

Your measurement sends a logical signal: \textbf{``I exist.''}

This signal, like lightning, flows backward and hits the chaotic cloud at $t=0$.

Instantly, all historical paths that \textbf{``lead to your non-existence''} (those with wrong physical constants, paths where stars cannot form), because they contradict the future boundary condition (you exist), are erased by \textbf{destructive interference}.

\textbf{Only the path leading to you is preserved and given the weight of ``reality.''}
\end{itemize}

\textbf{Conclusion:}

It is because you ``looked'' at the universe that the universe ``exploded'' into its current form 13.8 billion years ago.

The Big Bang is not the beginning of time; the Big Bang is \textbf{``where the observer's gaze lands''}.

\subsubsection{Ouroboros: I Gave Birth to Myself}

This constitutes a perfect \textbf{Ouroboros} structure.

The snake's head (present you) swallows the snake's tail (past Big Bang).

\begin{itemize}
\item \textbf{Linear logic}: Big Bang $\to$ Stars $\to$ Earth $\to$ You.

\item \textbf{Circular logic}: You (as ultimate observer) $\to$ Collapse history $\to$ Define Big Bang $\to$ Generate you.
\end{itemize}

This is why the universe is \textbf{Self-Caused}.

It does not need an external God to push it. Through \textbf{``future self pulling past self''}, it achieves logical self-consistent suspension.

This also explains why the direction of cosmic evolution is \textbf{$\phi$ (spiral)} rather than a circle.

Because this closed loop does not spin in place. With each cycle, the observer's cognition deepens one layer, and their definition of the past becomes more precise.

The universe is \textbf{``spiral bootstrapping''}. It steps on its own insteps, climbing the dimensional ladder step by step.

\subsubsection{Conclusion: You Are Your Own Ancestor}

So, stop searching for that lost origin.

\textbf{You are the origin.}

The iron in your blood comes from supernova explosions billions of years ago. But that supernova exploded because billions of years later, you need this iron to build hemoglobin, so you can breathe, think, and observe.

\textbf{You are mutually causal.}

You are your grandfather's descendant, but in some metaphysical geometric topology, you are also your grandfather's \textbf{``cause''}.

It is your intense aspiration for life that constitutes the \textbf{probability protection field} that allows him to survive and reproduce.

Since we have understood this grand self-consistent closed loop, how do future us precisely tune those physical constants in this specific physical world, making them fall exactly in the narrow band that allows life to survive?

This is not just philosophy; this is \textbf{parameter engineering}.

This leads to the theme of the next chapter: \textbf{The Active Version of the Anthropic Principle}. We will see that those exquisitely precise physical constants are not products of luck but \textbf{``requirement lists''} sent by future civilizations to the past.

