\subsection{6.1 Why Are Constants So Perfect?}

\begin{quote}
``If the universe is a suit, it fits us so well not because there are billions of suits in the store and we happened to pick this one, but because this suit is \textbf{custom-made}. The tailor is none other than the person wearing the suit---that is, our future selves.''
\end{quote}

\subsubsection{The Universe's Parameter Panel}

Imagine you stand before the control console of the Big Bang at $t=0$. Before you are dozens of knobs controlling the speed of light, charge strength, vacuum energy density, etc.

\begin{itemize}
\item If you turn the gravity knob up a bit, the universe will collapse back to a singularity seconds after the Big Bang.

\item If you turn the strong force knob down a bit, hydrogen cannot fuse into helium, and the universe will be a dark cloud of hydrogen gas.
\end{itemize}

It's as if someone at $t=0$ moment, with extreme precision, aligned all knobs to that extremely narrow frequency band (Goldilocks Zone) that \textbf{``allows intelligent life to emerge''}.

Who is this person? God?

No. In the causal closed loop, this person is the \textbf{observer at $\tau \gg 0$}.

\subsubsection{Retro-causal Parameter Locking}

In Volume V, we argued that observers collapse fuzzy history into definite trajectories through ``strong measurements.'' Now we apply this mechanism to \textbf{physical constants}.

In the quantum gravity era (Planck epoch), physical constants are not fixed values but \textbf{expectation values of Scalar Fields}. They are in \textbf{quantum superposition states}.

\[|\alpha\rangle = c_1 |\alpha=1/100\rangle + c_2 |\alpha=1/137\rangle + c_3 |\alpha=1/150\rangle + \dots\]

In the early evolution of the universe, all possible constant values were tried.

But \textbf{future boundary conditions} (i.e., ``must evolve intelligent life capable of observing the universe 13.8 billion years later'') impose a \textbf{``Post-selection''} filter on these superposition states.

\begin{itemize}
\item \textbf{Path A ($\alpha=1/100$)}: The evolved universe has no carbon, no life, no observers. $\implies$ No one observes in the future. $\implies$ The wave function of this path diverges due to lack of ``endpoint anchoring,'' amplitude goes to zero.

\item \textbf{Path B ($\alpha=1/137$)}: Carbon evolved, you were born. $\implies$ You observe in the future. $\implies$ This path forms a \textbf{self-consistent closed loop}, amplitude is resonantly amplified.
\end{itemize}

\textbf{Constants are perfect because only perfect constants can pass the ``existence test.''}

It's not that constants determine us, but \textbf{our existence reversely determines constants}.

\subsubsection{Projection of Requirements}

This is a \textbf{``requirement-oriented''} creation theory.

Future civilizations (high $v_{int}$ structures) impose strict \textbf{performance specifications} on the underlying hardware (physical laws):

\begin{enumerate}
\item \textbf{Stability requirement}: We need stable stars providing billions of years of energy $\implies$ Gravitational constant $G$ must be extremely small.

\item \textbf{Complexity requirement}: We need complex chemical reactions to build brains $\implies$ Fine structure constant $\alpha$ must exactly allow the existence of long-chain molecules.

\item \textbf{Computational requirement}: We need sufficient speed of light $c$ to support holographic network bandwidth.
\end{enumerate}

These requirements are encoded as the \textbf{``final state vector'' $|\Psi_{final}\rangle$}.

Through reverse propagation of path integrals, this final vector interferes with the initial vector $|\Psi_{initial}\rangle$, \textbf{``selecting''} the unique set of physical constants that can satisfy all requirements.

\subsubsection{Conclusion: There Is No Coincidence}

So, stop marveling at the coincidence of the universe.

Just as you don't need to marvel at why your shoes happen to fit your feet---because you bought them after trying them on.

\textbf{The universe is our ``custom exoskeleton.''}

It is the way it is because future us (that awakened, powerful observer) needs it to be this way, so we can run, think, and love within it.

We are not tenants of the universe; we are the \textbf{architects} of the universe. We stand at the end of the time axis, directing the foundation ceremony at the beginning of the time axis.

Since constants are determined by future requirements, this means we are participating in a \textbf{``engineering design''} across time.

Are our every wish now, every vision of the future, fine-tuning the underlying parameters of the universe?

This leads to the theme of the next section: \textbf{Participatory Creation}. We will see that we are not just passive beneficiaries but active \textbf{``parameter senders''}.

