\section{Nash Equilibrium of Objectivity}

\textbf{(客观性的纳什均衡)}

\begin{quote}
\textbf{"If everyone possesses an independent oracle (consciousness), if each observer can collapse wave functions to different branches through interaction, then why aren't we living in isolated illusion bubbles? Why is my 'red light' also your 'red light'? Objective reality is not some absolute truth; it is a Nash Equilibrium reached by countless interactive Turing machines in the Bekenstein game. Physical laws are this massive distributed consensus protocol."}
\end{quote}

In previous chapters, we established an \textbf{Interactive Computational Cosmology (ICC)} model based on a single observer. We proved that for a single player (CITM), the world is generated on demand. However, this model immediately faces the most severe challenge—the trap of \textbf{Solipsism}.

If the world is generated for you, then who am I? Am I a real NPC, or another player who also possesses an oracle interface? If we are all players with free will, when our wills conflict (I want the cat dead, you want it alive), whose should the system listen to?

This chapter will extend the ICC model from \textbf{Single-Player Mode} to \textbf{Multiplayer Mode}. We will prove that so-called "objective physical reality" is essentially a \textbf{Distributed State Synchronization} mechanism in multi-agent systems.

\subsection{Solipsism Dilemma and Many-Worlds Conflict}

In standard quantum mechanics, Wigner's Friend paradox reveals the contradiction of multiple observers:

\begin{itemize}
\item The friend measures spin in the laboratory and sees a definite "up."

\item Wigner outside the laboratory believes the friend is in a superposition of "up" and "down."
\end{itemize}

If there are $N$ observers, each trying to "instantiate" the world according to their own oracle input, the system faces the risk of \textbf{State Forking}. If the system allows everyone to have independent realities, the universe would split into mutually inaccessible private dreams, making scientific communication impossible.

Since we can converse, experiment, and reach agreement in a shared physical world, this indicates that the universe operating system runs a strict \textbf{Consensus Protocol}.

\subsection{Consensus Geometry: Reality as Intersection of Projections}

In the ICC model, we define "objective reality" as follows:

\begin{definition}[Objective Reality]
Objective reality is not the global Hilbert space $|\Psi\rangle$ (that's God's perspective), nor is it a single observer's private history $H_i$ (that's subjective perspective). Objective reality is the \textbf{Greatest Common Divisor} or \textbf{Intersection} of information within all local observers' horizons.

\begin{equation}
R_{obj} = \bigcap_{i=1}^N \text{View}_i
\end{equation}
\end{definition}

Imagine a massive multiplayer online game server.

\begin{itemize}
\item Player A sees a tree at a certain coordinate.

\item Player B also sees a tree at that coordinate.

\item To save resources, the system doesn't store two trees. The system maintains only one "tree" object in the backend database and broadcasts a reference to that object to both A and B simultaneously.
\end{itemize}

Therefore, \textbf{physical space is consensus space}. Only states that are jointly measured and locked by multiple observers have "hard" physical reality. States that exist only in one person's mind (such as hallucinations or private thoughts), due to lack of consensus signatures, are judged by the system as "illusory."

\subsection{Bayesian Updates and Wave Function Synchronization}

How is this consensus reached? Through \textbf{Bayesian Inference}.

Christopher Fuchs's \textbf{Quantum Bayesianism (QBism)} holds that wave functions are not objective entities, but observers' \textbf{Degree of Belief} about the future.

In multi-user systems, when two observers exchange information, their degrees of belief undergo \textbf{Synchronization}.

\textbf{Process Simulation}:

\begin{enumerate}
\item \textbf{Initial State}: Alice believes the electron is at A, Bob believes it's at B (belief conflict).

\item \textbf{Interaction}: Bob shouts to Alice: "I just measured it, it's at B!" (information exchange).

\item \textbf{Update}: Alice receives this bit stream. If she trusts Bob (regarding him as a reliable measuring instrument), she updates her prior probability according to Bayes' formula:

    \begin{equation}
    P(A|\text{Bob says B}) \to 0, \quad P(B|\text{Bob says B}) \to 1
    \end{equation}

\item \textbf{Consensus}: Now, both their wave functions collapse to B. Reality merges.
\end{enumerate}

\begin{theorem}[Consensus Convergence Theorem]
In an observer network with sufficient connectivity, as long as observers follow Bayesian rational update rules, their local wave functions will converge exponentially fast to a \textbf{globally consistent classical state}. This converged state is what we call "objective fact."
\end{theorem}

\subsection{Nash Equilibrium: Stability of Physical Laws}

Why does this consensus always converge to specific physical laws (such as $F=ma$) rather than magic or chaos?

This can be explained using \textbf{Nash Equilibrium} in game theory.

View the universe as a \textbf{Prediction Game}.

\begin{itemize}
\item Each observer's (player's) goal is to minimize their \textbf{Prediction Error} about the future (i.e., free energy principle).

\item Strategy: Players construct internal models (physical laws) to fit sensory inputs.
\end{itemize}

If everyone follows arbitrary rules, prediction errors will be large.

Only when everyone agrees on a set of \textbf{self-consistent, stable, universal} rules (standard model) does the total prediction error (information entropy) of the entire system reach a minimum.

\begin{corollary}[Physical Laws as Stable Strategies]
Newton's laws or quantum mechanics are not divine edicts carved in stone, but \textbf{Evolutionarily Stable Strategies (ESS)} in multi-agent systems.

\begin{itemize}
\item If I throw a stone, it falls down; you also see it fall down. This "gravity downward" model not only explains my data but also your data, and doesn't produce contradictions in interactions.

\item This model has \textbf{Robustness}, so it is preserved and solidified as a "law" by the system.
\end{itemize}
\end{corollary}

\subsection{The Distributed Ledger}

At the computer engineering level, this consensus mechanism is equivalent to the \textbf{Distributed Ledger} in blockchain technology.

\begin{enumerate}
\item \textbf{Block}: Every physical event (collapse) occurring at each Planck time step.

\item \textbf{Hash Chain}: Causality. Current state must contain encrypted signatures of past states, ensuring history is tamper-proof.

\item \textbf{Consensus Mechanism}:

    \begin{itemize}
    \item \textbf{Proof of Work (PoW)}: For macroscopic objects, changing their state requires consuming large amounts of energy (work). This prevents individual observers from arbitrarily modifying reality with thoughts.

    \item \textbf{Broadcast}: The speed of light $c$ is the maximum network latency for ledger synchronization. Once any event occurs, its effects broadcast to the entire universe at light speed. Once this information is recorded by enough nodes (environmental particles), the block is \textbf{Confirmed}, becoming irreversible objective history.
    \end{itemize}
\end{enumerate}

\textbf{Conclusion}:

We are not alone. Our consciousnesses are tightly connected through physical interaction networks. The so-called "objective world" is \textbf{a massive, decentralized, tamper-proof shared document} jointly maintained by all life forms (and measuring instruments). We are both readers and co-authors of this document.
