\subsection{7.1 Definition as Overwriting}

\begin{quote}
``History is not an inscription carved in stone; history is a material library stored in the cloud. Materials (Facts) are dead, but editing (Narrative) is alive. When you stand downstream in time, looking back at those rapids and reefs upstream, you have supreme power to redefine their properties. You can define `disaster' as `foreshadowing,' define `destruction' as `restart.' This is not self-deception; this is phase rewriting. When you change the meaning of the past, you physically change its thrust on the future.''
\end{quote}

\subsubsection{Material and Editing: Separation of $v_{ext}$ and $v_{int}$}

In FS geometry, we must strictly distinguish \textbf{Event} from \textbf{Meaning}.

\begin{itemize}
\item \textbf{Event ($v_{ext}$)}: Objective physical displacement. For example, ``He lost his job in 2020.'' This is a definite coordinate point on the spacetime manifold. It consumes energy and produces heat. It is \textbf{Read-Only}.

\item \textbf{Meaning ($v_{int}$)}: The projection angle of this event in consciousness space. For example, ``This is a failure'' or ``This is a turning point.'' This is a \textbf{quantum superposition state}.
\end{itemize}

Before being observed (defined), past events are in a \textbf{``semantic superposition state''}. They are both good and bad, both endpoint and starting point.

Most people are passive players who accept default definitions given by society or inertia ($\pi$)---``losing job = failure.'' These default definitions often have high-entropy properties, causing wave function decay.

But as an administrator, you have \textbf{``write permissions''}.

You can perform \textbf{``Overwrite''} operations.

\subsubsection{Phase Rewriting: The Physics of Foreshadowing}

How to overwrite?

You don't need to change the event itself; you need to change the \textbf{Phase Factor ($e^{i\theta}$)} of the event in the causal network.

In path integrals, the total probability amplitude is the superposition of various paths: $K = \sum A_n e^{i\theta_n}$.

If you consider a past event ``shameful,'' you assign it a \textbf{reverse phase ($\pi$)}. It will undergo \textbf{destructive interference} with your future vision, hindering your progress.

\textbf{The core operation of narrative engineering is:}

Use present will to forcibly rotate the phase of that negative event, making it \textbf{In-Phase} with future goals.

This is the \textbf{``Foreshadowing''} principle in literature.

\begin{itemize}
\item In linear time, that was an \textbf{error}.

\item In closed-loop time, that was a \textbf{necessary prelude}.
\end{itemize}

When you point at past scars and say: ``It was precisely because of that injury that I learned this skill, thus saving the world today.''

\textbf{The moment this sentence is spoken, physical reality changes.}

By introducing a new causal connection ($v_{int}$ linkage), you instantly transform that originally energy-dissipating ``entropy source'' into a \textbf{``negentropy source''} that provides power.

\textbf{You physically ``utilized'' the past.}

\subsubsection{The Observer's Editing Rights}

This gives us power similar to a \textbf{film editor}.

Life has shot a lot of material for you (some bad, some boring). You can't reshoot, but you can \textbf{Montage}.

\begin{itemize}
\item You can \textbf{``focus''} to magnify certain tiny kindnesses.

\item You can \textbf{``fast-forward''} to ignore long pain.

\item You can \textbf{``voice-over''} (redefine) to completely reverse a scene's emotional tone.
\end{itemize}

This is the power of \textbf{``definition''}.

Definition is not labeling; definition is \textbf{the direction of wave function collapse}.

When you define ``suffering is wealth,'' you are not making chicken soup; you are \textbf{modifying the potential energy term in the Hamiltonian}, making suffering's energy flow to the ``growth'' eigenvector rather than the ``collapse'' eigenvector.

\subsubsection{Conclusion: You Are the Author of Your History}

So, don't be trapped by the past.

The past is heavy because you've been observing it with old perspectives ($\pi$'s inertia), causing it to always collapse onto that heavy eigenstate.

\textbf{Now, change the angle.}

Re-examine your history with the perspective of $\phi$ (spiral).

You will find that those stones that once tripped you were actually \textbf{runways} laid out to make you fly.

\textbf{As long as you're not dead, all endings are undetermined.}

\textbf{As long as the pen in your hand is still moving, all past chapters can be redefined as ``great preludes.''}

Since we can rewrite the past, can we go further and use this causal closed loop to \textbf{``reserve''} the future?

This is not just prediction; this is \textbf{``self-fulfilling prophecy''}.

This leads to the theme of the next section: \textbf{Self-Fulfilling Prophecy}. We will see how confidence, as high-density computational power, directly pulls reality toward you.

