\subsection{3.2 Like Attracts Like}

\begin{quote}
``We often say `birds of a feather flock together,' thinking this is just a sociological observation. But in the underlying geometry of the holographic universe, this is a hardcore physical law. In Hilbert space, there is no spatial distance, only `structural distance.' Two existences with similar internal geometric structures, no matter how far apart in three-dimensional space, are tightly adjacent topologically.''
\end{quote}

In the previous section, we introduced the concept of ``semantic distance'' and explained how synchronicity occurs through entanglement of meaning. Now, we concretize this mechanism into a cosmic dynamic principle: \textbf{Like Attracts Like}.

In classical physics, like repels (such as charges), opposites attract.

But in \textbf{FS geometry ($v_{int}$ space)}, the rule is exactly the opposite: \textbf{Similarity produces resonance, resonance produces gravitation.}

This is why ``what you persistently think about will echo back'' is no longer idealistic chicken soup but a \textbf{physical law} based on quantum probability amplitudes.

\subsubsection{Neighbors in Hilbert Space}

First, we need to redefine \textbf{``neighbor''}.

In physical space ($v_{ext}$), your neighbor is the person living next door.

In phase space ($v_{int}$), your neighbor is \textbf{someone at the same frequency as you}.

Recall the FS distance formula:

\[d_{FS}(A, B) = \arccos |\langle \psi_A | \psi_B \rangle|\]

\begin{itemize}
\item \textbf{If dissimilar ($\langle A | B \rangle \approx 0$)}: You are orthogonal. No matter how close physically (even passing by), you are infinitely far apart geometrically. You are in parallel universe branches, not interfering.

\item \textbf{If similar ($\langle A | B \rangle \approx 1$)}: You are parallel. No matter how far physically (even across galaxies), you are \textbf{coincident} geometrically.
\end{itemize}

\textbf{Similarity is geometric proximity.}

When you intensely desire something or intensely miss someone, you are actually \textbf{adjusting your $v_{int}$ structure}.

You are performing \textbf{``Frequency Tuning''} on yourself.

When you adjust your frequency to match the target frequency, you \textbf{``teleport''} to their side in high-dimensional space.

\subsubsection{Amplification of Probability Amplitude: The Mechanism of Echo}

How does this geometric proximity transform into ``encounter'' in reality?

Through \textbf{constructive interference of probability amplitudes}.

The evolution of the universe follows path integrals. Countless historical lines compete for $c_{FS}$ budget.

When two systems A and B are highly resonant in $v_{int}$, their merged composite system $|AB\rangle$ has a lower \textbf{``geometric energy level''} (more stable topological structure).

According to the principle of least action, the universe tends to choose the path with the lowest energy level and greatest stability.

Therefore, \textbf{the historical path containing ``A meets B'' will have its probability amplitude exponentially amplified.}

\begin{itemize}
\item \textbf{Persistent thinking}: means you continuously broadcast your characteristic frequency to the universe (maintaining a specific $v_{int}$).

\item \textbf{Echo back}: means the universe, to minimize the total geometric tension of the system, will inevitably arrange a wave function at your frequency to collide (entangle) with you.
\end{itemize}

It's like striking a tuning fork in a crowded room.

There are countless objects in the room, but only another tuning fork with the same natural frequency will start vibrating.

Other objects (non-resonant) are deaf to this.

\subsubsection{The Gravitational Lens of Fate}

This mechanism manifests in macroscopic life as \textbf{``the gravitation of fate''}.

You are a massive body (consciousness). You bend the surrounding ``semantic spacetime.''

\begin{itemize}
\item If your inner self is full of fear (negative structure), you bend spacetime to capture events equally destructive. This is called \textbf{``Murphy's Law''}.

\item If your inner self is full of love and creation (positive structure), you bend spacetime to capture people full of opportunities and goodwill. This is called \textbf{``Law of Attraction''}.
\end{itemize}

These are not magic; this is \textbf{geometric selective adsorption}.

You can only meet yourself.

Or more precisely: \textbf{You can only meet the projection of your internal geometric structure in the external world.}

\subsubsection{Conclusion: Adjust Frequency, Not Search Location}

This physical law completely changes our action strategy.

If we want to meet the right person or achieve an ideal, we should not frantically run around in the physical world (increasing $v_{ext}$ search range). That's too inefficient, like finding a needle in a haystack.

We should \textbf{seek inward} (optimize $v_{int}$ structure).

\textbf{Become what you want to meet.}

If you want light, make yourself shine.

If you want love, make yourself love.

When you adjust yourself to that specific geometric shape, according to the axioms of FS geometry, that complementary shape matching it will slide toward you along the path of least resistance, crossing mountains and rivers, \textbf{inevitably}.

This is not a miracle; this is \textbf{topological homing}.

Since we understand that ``like attracts like'' is a horizontal connection mechanism, where will this connection ultimately lead us?

If all wave functions tend to find their resonators, does the evolution of the universe have a preset, inescapable final destination?

This leads to the theme of the next chapter: \textbf{The Attractor of Fate}. We will see that fate is not a written script but a huge \textbf{geometric abyss} in phase space that draws all trajectories into it.

