\chapter{Index of Symbols and Axioms}

This book constructs a vast theoretical system involving mathematical languages from multiple domains including quantum information, differential geometry, operator algebras, and complex systems. For readers' convenience in cross-referencing, this appendix compiles core axioms, main theorems, and definitions of key physical symbols throughout the book.

\section{Core Axiomatic System}

The theoretical edifice of this book is built on the following mutually supporting fundamental axioms. These axioms are not arbitrary assumptions, but minimal physical commitments made under the discrete ontology framework to reconcile contradictions between quantum mechanics and general relativity.

\begin{itemize}
\item \textbf{Axiom A1: Finite Information Density Axiom}

\begin{itemize}
\item \textbf{Definition}: Physical reality consists of discrete information units. For any finite spatial volume $V$, the number of independent physical degrees of freedom $N(V)$ it contains is finite, and there exists a natural cutoff at Planck scale.

\item \textbf{Source}: Chapter 1, Section 1.1.

\item \textbf{Corollary}: Hilbert space is locally finite-dimensional; continuum is only an effective approximation.
\end{itemize}

\item \textbf{Axiom A2: Finite Information Axiom}

\begin{itemize}
\item \textbf{Definition}: The Hilbert space $\mathcal{H}_{\text{phys}}$ of physical reality is isomorphic to a finite-dimensional vector space $\mathbb{C}^D$ over complex numbers on any compact spatial region.

\item \textbf{Source}: Chapter 1, Section 1.2.

\item \textbf{Corollary}: True infinities do not exist; UV divergences are naturally eliminated.
\end{itemize}

\item \textbf{Axiom A3: Causal Locality Axiom}

\begin{itemize}
\item \textbf{Definition}: The global update operator $U$ of QCA is composed of local rules; information propagation speed is limited by finite lattice step size (speed of light $c$).

\item \textbf{Source}: Chapter 3, Section 3.2.

\item \textbf{Corollary}: Light cone structure is strict; ``action at a distance'' is forbidden.
\end{itemize}

\item \textbf{Axiom A4: Maximum Entanglement Equilibrium Axiom (MEEA)}

\begin{itemize}
\item \textbf{Definition}: Under fixed causal geometric volume constraints, the generalized entropy $S_{\text{gen}}$ of vacuum state is at local stationary (extremal) state.

\item \textbf{Source}: Chapter 12, Section 12.1.

\item \textbf{Corollary}: Einstein field equations $G_{\mu\nu} = 8\pi G T_{\mu\nu}$ are the equation of state for spacetime thermodynamics.
\end{itemize}

\item \textbf{Axiom A5: Quantum Focusing Conjecture (QFC)}

\begin{itemize}
\item \textbf{Definition}: For any null geodesic congruence, the generalized expansion scalar $\Theta$ is monotonically decreasing along affine parameter: $d\Theta/d\lambda \le 0$.

\item \textbf{Source}: Chapter 14, Section 14.3.

\item \textbf{Corollary}: Ensures validity of generalized second law of thermodynamics (GSL) and stability of spacetime.
\end{itemize}

\item \textbf{Axiom A6: Topological Reality of Consciousness Axiom}

\begin{itemize}
\item \textbf{Definition}: True conscious agency is equivalent to the existence of a topologically protected $\mathbb{Z}_2$ Berry phase (holonomy index $\nu=1$) in the physical carrier.

\item \textbf{Source}: Chapter 28, Section 28.2.

\item \textbf{Corollary}: Consciousness is an objective physical state, not merely computational function.
\end{itemize}
\end{itemize}

\section{Physical Quantities and Symbol Index}

\subsection{Spacetime \& Geometry}

\begin{table}[h]
\centering
\small
\begin{tabular}{|p{2cm}|p{3cm}|p{5cm}|p{2cm}|}
\hline
\textbf{Symbol} & \textbf{Name} & \textbf{Definition/Physical Meaning} & \textbf{Source} \\
\hline
$g_{\mu\nu}$ & \textbf{Metric Tensor} & Describes curved structure of spacetime, arising from consensus geometry of information. & 2.1, 22.2 \\
\hline
$\mathcal{A}$ & \textbf{Unified Connection} & $\mathbb{A} = \frac{1}{2}\omega^{ab}J_{ab} + A^I T_I$, unifying gravity and gauge fields. & 16.2 \\
\hline
$\mathbb{F}$ & \textbf{Unified Curvature} & $\mathbb{F} = d\mathbb{A} + \mathbb{A}\wedge\mathbb{A}$, containing Riemann curvature and Yang-Mills field strength. & 16.3 \\
\hline
$D(p,q)$ & \textbf{Causal Diamond} & $J^+(p) \cap J^-(q)$, defining basic geometric unit of local thermodynamics. & 11.1 \\
\hline
$\theta$ & \textbf{Expansion Scalar} & Relative rate of change of cross-sectional area of geodesic congruence, describing gravitational focusing. & 11.4 \\
\hline
$\Lambda$ & \textbf{Cosmological Constant} & Lagrange multiplier maintaining holographic volume constraint of spacetime, arising from vacuum density of states. & 12.3 \\
\hline
\end{tabular}
\caption{Symbols for spacetime and geometry}
\end{table}

\subsection{Quantum \& Information}

\begin{table}[h]
\centering
\small
\begin{tabular}{|p{2cm}|p{3cm}|p{5cm}|p{2cm}|}
\hline
\textbf{Symbol} & \textbf{Name} & \textbf{Definition/Physical Meaning} & \textbf{Source} \\
\hline
$\rho$ & \textbf{Density Matrix} & Describes statistical state of quantum system, $\rho \in \mathcal{A}_{\text{int}}$. & 18.2 \\
\hline
$S_{\text{gen}}$ & \textbf{Generalized Entropy} & $S_{\text{gen}} = \frac{A}{4G} + S_{\text{mat}}$, core potential function of holographic gravity. & 11.2 \\
\hline
$\Phi$ & \textbf{Integrated Information} & Measures irreducible information flux of feedback closed loops in causal networks (IIT). & 20.3 \\
\hline
$\nu$ & \textbf{Topological Index} & $\mathbb{Z}_2$ holonomy index, distinguishing conscious state ($\nu=1$) from unconscious state ($\nu=0$). & 21.1 \\
\hline
$K$ & \textbf{Modular Hamiltonian} & $K = -\ln \rho$, generating energy flow in first law of entanglement. & 11.3 \\
\hline
$\mathcal{F}$ & \textbf{Variational Free Energy} & Functional measuring prediction error between internal model and external environment (FEP). & 19.3 \\
\hline
\end{tabular}
\caption{Symbols for quantum and information}
\end{table}

\subsection{Scattering \& Time}

\begin{table}[h]
\centering
\small
\begin{tabular}{|p{2cm}|p{3cm}|p{5cm}|p{2cm}|}
\hline
\textbf{Symbol} & \textbf{Name} & \textbf{Definition/Physical Meaning} & \textbf{Source} \\
\hline
$S(E)$ & \textbf{Scattering Matrix} & Unitary operator describing interaction processes, connecting incoming and outgoing states. & 6.2 \\
\hline
$\mathsf{Q}(E)$ & \textbf{EWS Operator} & $\mathsf{Q} = -i\hbar S^\dagger S'$, whose eigenvalues correspond to microscopic dwell times. & 6.2 \\
\hline
$\xi(E)$ & \textbf{Spectral Shift Function} & Describes change in level counting caused by interactions, $\det S = e^{-2\pi i \xi}$. & 7.1 \\
\hline
$\kappa(E)$ & \textbf{Unified Time Density} & $\kappa = \varphi'/\pi = \rho_{\text{rel}} = \text{Tr}\mathsf{Q}/2\pi$, microscopic master scale of time flow. & 8.1 \\
\hline
$\tau_W$ & \textbf{Wigner Delay} & Time delay experienced by wave packet in scattering region, macroscopically manifesting as gravitational redshift. & 6.3 \\
\hline
\end{tabular}
\caption{Symbols for scattering and time}
\end{table}

\section{Abbreviations}

\begin{itemize}
\item \textbf{QCA}: Quantum Cellular Automata

\item \textbf{IGVP}: Information Geometric Variational Principle

\item \textbf{MEEA}: Maximum Entanglement Equilibrium Axiom

\item \textbf{QNEC}: Quantum Null Energy Condition

\item \textbf{QFC}: Quantum Focusing Conjecture

\item \textbf{EWS}: Eisenbud-Wigner-Smith (time delay operator)

\item \textbf{MSCC}: Minimal Strongly Connected Component

\item \textbf{DTC}: Discrete Time Crystal

\item \textbf{SMC}: Symmetric Monoidal Category

\item \textbf{DCC}: Dagger Compact Category

\item \textbf{PSWF}: Prolate Spheroidal Wave Functions

\item \textbf{DCEC}: Discrete-Continuous Error Control
\end{itemize}

\textbf{Acknowledgments and Copyright Notice}

This book is an axiomatic reconstruction attempt based on current theoretical physics frontiers (holographic principle, quantum information, category theory). All cited theorems and experimental data are based on publicly available scientific literature (see references in each chapter). The unified framework (QCA+IGVP+IIT) proposed in this book aims to provide a new perspective for future physics research.

\textbf{(End of Appendices)}

