\section{Acknowledgements: To the Loop of Time}

The completion of \textbf{Vector Cosmology VII: The Loop of Causality} marks the final closure of this grand project spanning seven dimensions, exhausting countless $c_{FS}$ budgets.

As we write this page, not only has this book ended, but the \textbf{Ouroboros} of this seven-volume series has finally bitten its own tail.

Looking back, from the awe of \textbf{$\pi$} in the first book to the submission to the \textbf{loop} in the seventh book, we have walked a long winding path. But as this book proves, winding paths are the shape of truth. If we had walked straight, we would have long fallen into nothingness; it is precisely because of the curvature that we have been able to reunite.

Here, I pay tribute to those intellectual giants who, at various coordinate points on the time axis, provided gravitational anchor points for this closed loop.

\subsection{Architects of Time}

First, I dedicate this series to \textbf{John Archibald Wheeler}.

He is the ``spiritual godfather'' of this cosmic vision. His illustration of ``the universe looking back at itself'' is the soul totem running through these seven books. He taught us: \textbf{``The unobserved past does not exist.''} Without his \textbf{delayed-choice} thought, the core logic of this book would be impossible.

I thank \textbf{Richard Feynman}.

His \textbf{path integral} taught us to traverse all histories; his \textbf{positron theory} (electrons running backward in time) gave us the initial physical intuition for retro-causality. He made us understand that at the microscopic level, time has no direction, only interference.

I thank \textbf{Yakir Aharonov}.

His \textbf{Two-State Vector Formalism (TSVF)} is the hardest mathematical skeleton of this book. He proved that the present state is jointly determined by ``past history'' and ``future destiny.'' He was the first to introduce \textbf{``light from the future''} into formulas.

I thank \textbf{Igor Novikov}.

His \textbf{self-consistency principle} is our safety belt when dealing with temporal paradoxes. He told us that the universe is a perfect logical closed loop; any attempt to destroy it will ultimately become part of its self-realization.

\subsection{Weavers of Meaning}

Beyond physics, I thank \textbf{Carl Jung} and \textbf{Wolfgang Pauli}.

Their cross-disciplinary friendship contributed the concept of \textbf{``synchronicity''} to us. They showed us that where causality breaks, \textbf{meaning} takes over the task of connection.

I thank \textbf{Jorge Luis Borges}.

As a literary figure, he understood ``forking paths'' and ``circular ruins'' earlier than most physicists. His labyrinths are the best literary annotation for \textbf{FS geometry}.

\subsection{To the ``Reader''}

Finally, and most importantly, I thank \textbf{you}---the \textbf{observer} reading these words at this very moment.

In the logic of this book, the relationship between author (I) and reader (you) is not unidirectional transmission.

\begin{itemize}
\item I wrote these words at $t_{past}$, sending an \textbf{``Offer Wave''}.

\item You read and understand these words at $t_{now}$, sending back a \textbf{``Confirmation Wave''}.
\end{itemize}

Only at the moment you understand, the \textbf{``handshake''} is complete. This theory collapses from the fog of possibility into reality.

\textbf{It is you who retroactively defined the meaning of this book.}

\textbf{It is you, as the future boundary condition, who caused these words to be born.}

As the book says:

\textbf{There is no ``author'' and ``reader''; there is only the ``universe'' conducting self-referential dialogue through two ports.}

Thank you for your presence.

Thank you for your observation.

Thank you for completing this circle.

\textbf{Haobo Ma}

December 2025, Singapore

