\section{Glossary: The Lexicon of the Loopers}

\textbf{Vector Cosmology VII: The Loop of Causality} pushes the exploration of physics to the limits of logic. It breaks the shackles of linear time, revealing a cosmic picture that is end-to-end connected and self-generating. To help readers understand this ultimate model filled with retro-causality, synchronicity, and self-referential structures, this glossary redefines the core concepts in the book from the perspective of the ``closed loop.''

\subsection{Time \& Causality}

\subsubsection{Ouroboros}

The ultimate topological structure of the universe. Symbolizes the closure of causal chains: the endpoint ($\Omega$) is not only the evolutionary result of the starting point ($t=0$), but also the \textbf{``boundary condition''} of the starting point. For the universe to exist, it must devour and generate itself.

\subsubsection{Retro-causality}

A corollary of temporal symmetry in FS geometry. Refers to the fact that future states (observer's measurement choices or existence itself) can retroactively determine past historical paths. History is not an established fact, but a \textbf{``pending draft''} awaiting future ``confirmation.''

\subsubsection{Delayed Choice}

A core mechanism of quantum mechanics. Proves that a photon's path choice in the past (wave or particle) depends on the current observation method. Macroscopically, this means observers possess the privilege of \textbf{``retroactive historical revision''}; present cognition can reshape the meaning of the past.

\subsubsection{Quantum Handshake}

A reality-generating mechanism based on transactional interpretation. Physical reality results from the meeting of a \textbf{``retarded wave''} (from the past) and an \textbf{``advanced wave''} (from the future) in spacetime, forming a standing wave.

\subsection{Meaning \& Connection}

\subsubsection{Synchronicity}

Non-causal \textbf{``semantic entanglement''}. When two independent systems are highly similar in $v_{int}$ (internal structure), their distance in semantic space approaches zero, leading to \textbf{``coincidental''} events in the physical world. It is the overlap of high-dimensional geometry in low-dimensional projections.

\subsubsection{Like Attracts Like}

The selective adsorption law in Hilbert space. Wave functions with similar frequencies or geometric configurations tend to resonate and establish entanglement. This is the physical foundation of fate's gravity.

\subsubsection{Semantic Wavefunction}

A quantum state describing the meaning structure within a conscious entity. It evolves in phase space, and its phase interference determines the fortune or misfortune of personal destiny (constructive or destructive).

\subsubsection{Attractor of Fate}

A geometric steady state in phase space. Regardless of initial conditions, the evolutionary trajectory of a dynamical system will eventually converge to a specific region. It is the \textbf{``destined outcome''} preset by physical laws, but free will can fine-tune the \textbf{``posture''} of entering that outcome.

\subsection{Creation \& Structure}

\subsubsection{Bootstrap}

The self-consistent creation logic of the universe. The universe evolves future observers, who retroactively set the initial parameters of the Big Bang, thus achieving \textbf{``self-creation''}. This is a logical infinite loop without external input, but geometrically a perfect self-consistent closed loop.

\subsubsection{Active Anthropic Principle}

An explanation for the fine-tuning of physical constants. It is not that we adapted to the universe, but that future us sent a \textbf{``system requirements list''} to the past, forcing the universe to generate constants suitable for life (such as $G, \alpha, c$) at the Big Bang.

\subsubsection{Narrative Engineering}

A technique for observers to edit personal timelines using retro-causality. By changing the \textbf{``definition''} (phase rotation) and \textbf{``weight''} (attention renormalization) of past events, observers can transform negative history (entropy source) into positive momentum (negentropy source).

\subsubsection{Quantum Erasure}

A physical operation that modifies historical information. By eliminating information about ``paths'' (such as forgiveness or letting go), the interference fringes of the wave function can be restored, returning collapsed trauma states to flowing superposition states.

\subsection{Ontology \& Ethics}

\subsubsection{Mirror Theory}

The ultimate relationship between self and others. All conscious entities are projections of the same global vector onto different orthogonal bases. Others are selves in different coordinate systems. Harming others is self-harm; loving others is self-healing.

\subsubsection{Indra's Net}

A holographic mutual-reflection network topology. There are no isolated entities in the universe; every node (consciousness) reflects the radiance of all other nodes. Objective reality is a \textbf{``consensus eigenstate''} formed by recursive reflection of countless subjective dreams.

\subsubsection{The Ultimate Sufferer}

An ethical corollary based on the single-player model. At the macroscopic scale, the universe has only one experiencer (noumenon). All pain and evil are ultimately borne by this same noumenon.

\subsubsection{Return to One}

The endpoint of the seven-volume series. Refers to observers, after understanding circle, spiral, generator, light speed, value, space, and causality, finally breaking the ``subject-object dualism'' and geometrically returning to that state of \textbf{``undifferentiated Oneness''}.

