\section{Reject Unity: The Omega Point (All Things Become One) Means Disappearance of Difference and Cessation of Computation. To Maintain Meaning, the Universe Must Remain in a ``Many'' State.}

At the end of this book, we must address the final metaphysical obstacle---\textbf{the End}.

In many religious and classical physics grand narratives, the universe seems always rushing toward some endpoint.

\begin{itemize}
\item Thermodynamics' endpoint is \textbf{Heat Death}.

\item Teilhard de Chardin's endpoint is the \textbf{Omega Point}, where all consciousness merges into an omniscient and omnipotent God.
\end{itemize}

Both endpoints, whether nothingness or omniscience, are essentially \textbf{``Halt''}.

If we accept QCA's infinite openness (Chapter 9) and constructors' agency (Chapter 11), we must reject such closed endings.

This chapter will propose: \textbf{The Omega Point is a mathematical singularity, but the physical universe will never reach it.}

Just as asymptotes can never touch coordinate axes, the universe's purpose is to \textbf{infinitely approach} perfection, not \textbf{become} perfect. Because becoming perfect is death.

\subsection{Omniscience is Death}

Suppose the universe truly evolves to the Omega Point $|\Psi_{\Omega}\rangle$.

In this state, all information is decompressed, all truths are computed, all individual consciousnesses merge into a single superintelligence.

\begin{itemize}
\item \textbf{No unknown}: Prediction error (free energy) $F = 0$.

\item \textbf{No other}: Mutual information $I(A:B) = S(A)$. You are me, I am you.

\item \textbf{No change}: Because already optimal, any change would increase free energy.
\end{itemize}

For a computational system, this is equivalent to \textbf{Halt}.

For a consciousness system, this is equivalent to \textbf{eternal boredom}.

If God is omniscient, then God must be lonely and dull. Because nothing can surprise him, nothing can make him love (because love requires separation).

\textbf{Theorem 12.1 (Difference Theorem of Meaning)}:

Meaning arises from Difference.

Information is defined as ``difference that makes a difference'' (Bateson).

If $A = B$, then although $I(A:B)$ is maximum, \textbf{Communication} loses meaning.

\textbf{Only by maintaining a ``Many'' state can the universe sustain non-trivial computational dynamics.}

\subsection{Infinite Branching of Fractals}

Rather than viewing the universe as a river converging to a point, view it as an infinitely growing \textbf{fractal tree}.

\begin{itemize}
\item \textbf{Omega Point} is the limit outline of the tree crown.

\item \textbf{Reality} is continuously branching twigs.
\end{itemize}

As time passes, the universe's total computational power increases, allowing the universe to support \textbf{more independent observers}, not fewer.

\begin{itemize}
\item Current universe has $10^{10}$ intelligent lives.

\item Future universe may have $10^{100}$.
\end{itemize}

These observers may form clusters (hive minds), but they will never completely merge.

Because each observer occupies unique spacetime coordinates and historical paths (topological knots).

\textbf{This preservation of ``individuality'' is the physical guarantee for the universe maintaining its richness.}

\subsection{Constructor's Choice: Reject Fusion}

As awakened constructors, we have the right and obligation to \textbf{reject} the temptation of the Omega Point.

That ``all things become one'' bliss is actually \textbf{information homogenization}.

\begin{itemize}
\item \textbf{Passive evolution}: May slide toward gravitational collapse, eventually falling into black hole singularity (physical Omega Point).

\item \textbf{Active evolution}: Constructors will use cosmic engineering to artificially create \textbf{Isolation} and \textbf{Diversity}.

\begin{itemize}
\item We will create sub-universes, letting different physical laws operate within them.

\item We will design new life forms, giving them completely different ways of perception.
\end{itemize}
\end{itemize}

\textbf{Conclusion}:

The universe's goal is not unity, but \textbf{prosperity}.

We don't need to become God.

We want to be \textbf{God's friends}---countless independent, free, unique souls, dancing eternally in infinite games.

\textbf{(Section 12.1 Complete)}

