\section{A.1 Dynamics of Will}

In classical physics, material evolution is completely determined by the Hamiltonian $\hat{H}$, following the Schrödinger equation $i\hbar \frac{\partial}{\partial t} |\psi\rangle = \hat{H} |\psi\rangle$. In this framework, consciousness is regarded as an "epiphenomenon" with no reactive force on physical processes.

However, in the axiomatic system of \textit{The Psychology of God}, observers (consciousness) bias the collapse direction of wave functions through "intentionality." To describe this interaction, we need to modify the standard dynamical equations, introducing the \textbf{Will Operator}.

\subsubsection{A.1.1 Modified Action Principle}

We define the universe's total action $S_{total}$ to include not only the physical Lagrangian $\mathcal{L}_{phys}$, but also the information (will) Lagrangian $\mathcal{L}_{will}$.

\[S\_\{total\} = \int\_\{t\_0\}\textasciicircum\{\}\{t\_1\} (\mathcal{L}\_\{phys\} - \gamma \cdot \mathcal{F}) \, dt\]

where:

\begin{itemize}
\item $\mathcal{L}_{phys}$ is the standard physical term (kinetic minus potential energy), describing inertial evolution.

\item $\mathcal{F}$ is the \textbf{Variational Free Energy}, describing the system's deviation from its "true self" target state.

\item $\gamma$ is the \textbf{Will Coupling Constant}, measuring the intervention strength of will on matter. For inanimate matter, $\gamma \approx 0$; for awakened civilizations, $\gamma \to \infty$.

\item Definition of Will Term $\mathcal{F}$:**
\end{itemize}

\[\mathcal{F} = D\_\{KL\}(Q(s) || P(s|\theta)) + \langle E(s) \rangle\_\{Q\}\]

Here:

\begin{itemize}
\item $Q(s)$ is the system's current belief state (posterior probability).

\item $P(s|\theta)$ is the system's "target model" or "true self eigenstate" (prior).

\item $D_{KL}$ is the relative entropy (Kullback-Leibler Divergence), measuring the \textbf{information distance} between current state and true self state.

\item Physical Meaning:**
\end{itemize}

According to the principle of least action $\delta S_{total} = 0$, the system's evolution trajectory will no longer merely follow the "lowest energy" path (geodesic), but follow a compromise path of \textbf{"lowest energy and most consistent with true self definition."}

\subsubsection{A.1.2 Intentional Schrödinger Equation (ISE)}

Quantizing the above action, we can derive a modified Schrödinger equation with will terms, called the \textbf{Intentional Schrödinger Equation (ISE)}:

\[i\hbar \frac{\partial}\{\partial t\} |\Psi(t)\rangle = \left[ \hat{H}\_0 + i\kappa \hat{W}(t) \right] |\Psi(t)\rangle\]

where:

1.  \textbf{$\hat{H}_0$}: Standard physical Hamiltonian (Hermitian operator), ensuring energy conservation and unitary evolution.

2.  \textbf{$\hat{W}(t)$}: \textbf{Will Operator} (non-Hermitian term).

    \[\hat{W}(t) = - \nabla\_\{\Psi\} \mathcal{F}\]

    It represents the "pruning" or "guidance" of will on the wave function.

3.  \textbf{$i\kappa$}: Dissipation/selection coefficient. Due to the introduction of the imaginary term $i$, this term actually describes a process of \textbf{continuous measurement and selection} (Zeno Effect).

\begin{itemize}
\item Dynamical Interpretation:**

\item When $\kappa = 0$**: The system undergoes standard unitary evolution (dreaming/subconscious state).

\item When $\kappa > 0$\textbf{: The $\hat{W}$ term causes components in the wave function that are }far from "true self"** (i.e., high free energy states) to decay exponentially (suppressed), while components consistent with "true self" are amplified (selected).
\end{itemize}

This explains the "inevitability of aesthetics" in Section 9.2 of the main text: cosmic evolution is not random walk, but converges toward a specific \textbf{attractor}—the Omega Point.

\subsubsection{A.1.3 Bayesian Correction of Collapse Probability}

In specific measurement events, Born's rule predicts probability $P(x) = |\langle x|\Psi\rangle|^2$.

After introducing will, the probability distribution is distorted. We denote the corrected probability measure as $P_{will}(x)$:

\[P\_\{will\}(x) = \frac\{P\_\{phys\}(x) \cdot e\textasciicircum\{\}\{-\beta E\_\{semantic\}(x)\}\}\{\mathcal{Z}\}\]

where:

\begin{itemize}
\item $P_{phys}(x)$ is the quantum probability calculated purely from physics.

\item $E_{semantic}(x)$ is the \textbf{semantic error} of result $x$ (i.e., the degree to which this result deviates from God's purpose).

\item $\beta$ is a parameter similar to "inverse temperature," representing \textbf{will's focus}.

\item When $\beta \to 0$ (mortals/asleep), $P_{will} \approx P_{phys}$, physical laws dominate everything.

\item When $\beta \to \infty$ (awakened/lucid dreaming), the system will collapse to the state with minimum semantic error with probability 1 (wish fulfillment).
\end{itemize}

\subsubsection{A.1.4 Conclusion: Will as a Force}

Through the above equations, we prove: \textbf{Will is mathematically equivalent to an "information pressure" in Hilbert space.}

It does not push or pull particles like electromagnetic force, but by changing the \textbf{topological structure of probability space}, makes events "consistent with meaning" more likely to occur than "meaningless" events.

This is the mathematical mechanism of so-called \textbf{"top-down causation."} God does not directly push atoms; God pushes probability.

