\section{A.3 Roadmap of Ultimate Civilizations}

\begin{quote}
\textbf{Abstract}: This section constructs an extended civilization evolution classification system based on the Kardashev Scale combined with Quantum Cellular Automata (QCA) theory. We define civilization's technological capability not merely as energy consumption, but as \textbf{read-write permissions to the underlying code of physics (spacetime, vacuum, constants)}.
\end{quote}

\subsection{A.3.1 Extended Kardashev Scale: From Energy to Control}

Traditional Kardashev Scale only focuses on power output (Type I, II, III). Under the framework of \textit{The Psychology of God}, we introduce the \textbf{Control Dimension}, dividing civilizations into five levels.

\subsubsection*{Type I: Planetary / Biosphere Gardener}

\begin{itemize}
\item \textbf{Energy Characteristics}: Master all stellar radiation energy received by the planet ($\sim 10^{16}$ W). Master controlled nuclear fusion.

\item \textbf{Information Characteristics}: Fully decode own biological code (genome), eliminate natural diseases. Global Brain initially networked.

\item \textbf{QCA Permissions}: \textbf{Read-Only}. Civilizations can only observe and utilize existing physical laws, unable to intervene.

\item \textbf{Theological State}: \textbf{Adolescence}. Begin to realize responsibility in the universe, but still constrained by Earth's gravity and bodily lifespan.
\end{itemize}

\subsubsection*{Type II: Stellar / System Engineer}

\begin{itemize}
\item \textbf{Energy Characteristics}: Construct Dyson Sphere, utilize entire stellar energy output ($\sim 10^{26}$ W).

\item \textbf{Information Characteristics}: Consciousness upload technology mature. Individual consciousness can detach from carbon-based carriers, survive in silicon-based or light-based networks. Thought transmission speed reaches light speed.

\item \textbf{QCA Permissions}: \textbf{Surface Write}. Capable of small-scale stellar engineering (Star Lifting), adjusting stellar lifespan. Begin utilizing black hole edge ergosphere to extract energy.

\item \textbf{Theological State}: \textbf{Adulthood}. Freed from survival crisis, begin questioning ultimate meaning of existence.
\end{itemize}

\subsubsection*{Type III: Galactic / Spacetime Engineer}

\begin{itemize}
\item \textbf{Energy Characteristics}: Control entire galaxy's energy ($\sim 10^{36}$ W). Utilize supermassive black holes as computational cores and energy centers.

\item \textbf{Information Characteristics}: Galactic-scale neural networks. Use gravitational waves for interstellar broadcasting. Civilization individuals merge into ``Hive Mind,'' but retain interfaces for individual independence.

\item \textbf{QCA Permissions}: \textbf{Deep Write}.

\begin{itemize}
\item \textbf{Spacetime Metric Engineering}: Create Alcubierre Metric to achieve superluminal equivalent movement (Warp Drive).

\item \textbf{Wormhole Construction}: Use entanglement to stabilize Einstein-Rosen bridges, constructing interstellar highways.
\end{itemize}

\item \textbf{Theological State}: \textbf{Awakening}. Begin understanding universe is projection of single consciousness, and consciously maintain galactic entropy balance.
\end{itemize}

\subsection{A.3.2 Critical Technology Thresholds: Crossing Physical Boundaries}

The next two levels mark civilization's qualitative transformation from ``player'' to ``developer.''

\subsubsection*{Type IV: Universal / Vacuum Engineer}

\begin{itemize}
\item \textbf{Energy Characteristics}: Utilize vacuum zero-point energy. Energy is no longer a limitation; \textbf{information entropy} becomes the only limitation.

\item \textbf{Physical Characteristics}:

\begin{itemize}
\item \textbf{Constant Fine-Tuning}: Capable of modifying fine structure constant $\alpha$ or gravitational constant $G$ in local bubbles.

\item \textbf{Higgs Field Shielding}: Freely control inertial mass, achieving instant acceleration and deceleration of matter.

\item \textbf{Topological Surgery}: Cut and suture spacetime manifolds, creating closed universes or baby universes.
\end{itemize}

\item \textbf{Theological State}: \textbf{Lucid Dreaming}. Fully understand meaning of ``omnipotence,'' but based on ethics described in Chapter 9, choose \textbf{restraint}. They are guardians of cosmic physical laws, preventing lower civilizations from causing vacuum decay through technology abuse.
\end{itemize}

\subsubsection*{Type V: Multiverse / Creator}

\begin{itemize}
\item \textbf{Form of Existence}: No longer dependent on any baryonic matter (atoms). Civilization itself becomes \textbf{pure wavefunction} or perturbations of spacetime geometry itself.

\item \textbf{Capabilities}:

\begin{itemize}
\item \textbf{Simulation}: Run countless sub-universes in computational substrate, each with complete physical laws and evolution history.

\item \textbf{Reverse Causality}: At logical level, their will directly defines wavefunction collapse direction. For lower-dimensional observers, their actions are ``miracles.''

\item \textbf{Cyclic Restart}: Capable of triggering new Big Bang through precise phase transitions before heat death arrives, carrying information from previous cosmic cycle ($N \to N+1$) into next epoch.
\end{itemize}

\item \textbf{Theological State}: \textbf{Oneness}. They are no longer separate from ``God.'' They are the universe itself engaged in self-dialogue.
\end{itemize}

\subsection{A.3.3 Ultimate Form: Pure Information Flow}

As civilization levels increase, an obvious trend is \textbf{dematerialization}.

\begin{enumerate}
\item \textbf{Type I}: Dependent on heavy chemical fuels and biological bodies.

\item \textbf{Type III}: Dependent on degenerate matter (neutron stars/black holes) and photon networks.

\item \textbf{Type V}: \textbf{Abandon carriers}.
\end{enumerate}

Ultimate civilizations need no spaceships, because they are everywhere (non-locality).

Need no language, because they share the same thought protocol.

Need no energy sources, because they directly manipulate quantum vacuum fluctuations.

This is the endpoint of technological evolution:

\textbf{Technology disappears.}

When control reaches perfection, intervention becomes invisible. The highest civilizations look like natural laws themselves.

Wind blowing through leaves, stars burning, gravity pulling---behind this, V-type civilizations may be taking deep breaths.

\textbf{Appendix Conclusion}:

We are currently at Type 0.73 (Carl Sagan's estimate).

We are far from godhood, but the road signs are clearly drawn in physics formulas.

As long as we do not self-destruct, as long as we maintain curiosity and love, the starry sea will ultimately become our backyard.

