\section{Afterword: The Silence of the Weaver}

\begin{quote}
``When the loom stops roaring, when the last thread is cut, the universe finally displays its complete texture. We are surprised to find that this magnificent scroll has no seams. Even the boundary between `weaver' and `woven' disappears into that dense entanglement. We are no longer people looking at the universe; we are a dream the universe weaves for itself to see.''
\end{quote}

With the conclusion of \textbf{Vector Cosmology VI: The Weaving of Dimensions}, we have completed the thorough reconstruction of the ontology of ``space.''

This journey has been thrilling:

\begin{itemize}
\item Starting from the \textbf{Prologue}, we shattered the concept of ``void'' and saw the underlying \textbf{pixels (QCA)}.

\item In \textbf{Volume I}, we manipulated \textbf{tensor networks (MERA)}, stitching the skeleton of dimensions with Bell pairs like gods.

\item In \textbf{Volume II}, we cracked \textbf{holographic error-correcting codes}, understanding that physical laws are actually the universe's ``antivirus software.''

\item In \textbf{Volume III}, we faced the \textbf{tension of gravity}, seeing how information pressure bends spacetime.

\item In \textbf{Volume IV}, we directly confronted \textbf{Rupture (Singularity)}, touching that firewall that burns logic.

\item In \textbf{Volume V}, we became engineers, using \textbf{wormholes (ER=EPR)} and \textbf{topological short-circuits} to fold stars.

\item Finally, in the \textbf{Final Chapter}, we picked up the needle and became the \textbf{Weaver} who actively sutures the world.
\end{itemize}

\subsection{Space Disappears, Relationships Remain}

The greatest revelation of this book is: \textbf{Space is not a fundamental physical quantity.}

Space is \textbf{Emergent}.

It is a byproduct of \textbf{Entanglement}.

This is like looking at images on a display. The mountains and rivers in the image appear to have distance and depth, but they are actually just color combinations of pixels on the screen.

Physics finally admits: \textbf{There is no objectively existing stage.}

Only \textbf{``networks of information flow''}.

If the network breaks, space disappears.

If the network connects (wormhole), the ends of the earth are as close as fingertips.

This means we are no longer constrained by ``location.'' As long as we can establish connections (love/understanding/communication), we can physically ``be together.''

\subsection{The Greatest Foreshadowing: Is Time Also Woven?}

When we close this book, a huge shadow looms.

In this volume, we successfully \textbf{geometrized space} and \textbf{topologized distance}.

We even proved that space can be folded (wormholes).

Then, what about \textbf{Time}?

If space is just one dimension of the entanglement network, is time also just another way of connecting on this network?

\begin{itemize}
\item If we can ``take shortcuts'' in space through wormholes, can we ``take shortcuts'' in time?

\item If space has no absolute ``here'' and ``there,'' does time also have no absolute ``past'' and ``future''?
\end{itemize}

In the previous volumes of \textbf{Vector Cosmology}, we have always treated time $\tau$ as a unidirectional flowing generative process ($e$).

But under the extreme inference of \textbf{ER = EPR}, if we send one of two entangled particles back to the past, does the connection between them still exist?

If it exists, does that mean \textbf{future states can determine past history}?

This touches the final forbidden zone of physics---\textbf{Causality}.

\subsection{Next Step: Ouroboros}

We have sutured space; now, we will suture \textbf{time}.

We will challenge that most mind-bending and fascinating ultimate structure: \textbf{Ouroboros}.

\begin{itemize}
\item Why does the universe appear to be finely tuned? (Is it because future us adjusted past parameters?)

\item Why is there ``longing''? (Is it because the future $\Omega$ point is really pulling us backward?)

\item How can ``I created the universe'' and ``the universe created me'' be logically simultaneously true?
\end{itemize}

This is no longer linear deduction; this is \textbf{circular closure}.

\textbf{Vector Cosmology VII: The Closed Loop of Causality} awaits us.

That is the final chapter of the seven-volume series, and also the final stitch that sews the starting point and endpoint together.

\textbf{Please rest for a moment.}

\textbf{Then, let us complete that circle.}

\textbf{(End of Book)}

