\section{Glossary: The Lexicon of the Weavers}

\textbf{Vector Cosmology VI: The Weaving of Dimensions} fundamentally subverts the classical ``container space view'' and establishes a new physical picture of ``space as an entanglement network.'' To help readers understand this complex geometric world constructed from tensors, qubits, and holographic duality, this glossary redefines the core concepts in the book from the perspective of ``the weaver.''

\subsection{1. Network \& Geometry}

\subsubsection{Tensor Network}

The microscopic skeleton of cosmic spacetime. It is not a continuous manifold, but a graph formed by countless discrete quantum logic gates (tensors) connected through entanglement. The space we perceive macroscopically is the emergence of this network in the long-wavelength limit.

\subsubsection{Bell Pair}

The smallest stitching unit of space. Two qubits in a maximally entangled state ($|\Phi^+\rangle$). Geometrically, each Bell pair is a microscopic thread that ``pulls'' two separated points together. Cutting a Bell pair equals cutting space.

\subsubsection{MERA (Multiscale Entanglement Renormalization Ansatz)}

A specific tensor network architecture of spacetime. Through layers of stacked ``disentanglers'' and ``isometries,'' it constructs a fractal tree. The horizontal connections of this tree generate 3D space, while the vertical layers generate ``energy scales'' or ``holographic depth'' ($z$-axis).

\subsubsection{Entanglement Entropy}

A physical quantity measuring the strength of spatial connections. According to the Ryu-Takayanagi formula ($S = A/4G$), the entanglement entropy of a region strictly equals the minimum cross-sectional area required to cut it off from its environment. Entropy is geometric area.

\subsection{2. Holography \& Code}

\subsubsection{Holographic Principle}

The information-theoretic axiom of the universe. It asserts that all information within a spatial volume ($v_{int}$ structure) can be losslessly encoded on its lower-dimensional boundary. We live in the holographic projection of the ``bulk,'' while our true selves reside on the distant ``boundary.''

\subsubsection{AdS/CFT Correspondence}

The concrete mathematical realization of the holographic principle. It is a dictionary that translates ``quantum field theory code on the boundary'' into ``curved spacetime geometry in the bulk.'' Gravity is not a fundamental force, but the geometric manifestation of boundary code complexity.

\subsubsection{Holographic Error Correction}

The essence of physical laws. To prevent fragile microscopic quantum states from being destroyed by environmental thermal noise, the universe ``smears'' logical information across the entire holographic boundary. The robustness of spacetime stems from this redundant encoding; conservation laws are the checksum operators of the code.

\subsubsection{Holographic Complexity}

A physical quantity corresponding to the internal volume of black holes. It measures the total number of logic gate operations required to generate the current quantum state. Spatial volume is the ``fossil'' of the universe computer's operational history.

\subsection{3. Gravity \& Topology}

\subsubsection{Entropic Gravity}

A thermodynamic interpretation of gravity. When matter occupies the bandwidth of the tensor network and reduces local entanglement entropy, spacetime fabric produces an elastic recoil force attempting to restore maximum entropy. Spacetime curvature is information congestion.

\subsubsection{ER = EPR}

A bridge connecting quantum mechanics and general relativity. The Einstein-Rosen bridge (wormhole, ER) is mathematically equivalent to the Einstein-Podolsky-Rosen pair (entanglement, EPR). Microscopic entanglement is a microscopic wormhole; macroscopic entanglement is an interstellar stargate.

\subsubsection{Firewall}

A rupture surface of spacetime continuity. When the monogamy of black hole radiation entanglement is violated, to maintain unitarity, spatial connections at the horizon are forcibly severed, forming a wall of high-energy particles. It is the physical manifestation of logical contradictions in physical laws.

\subsubsection{Topological Short-Circuit}

An interstellar travel technology based on ER=EPR. By creating macroscopic-scale entanglement, direct channels are established at the folds of spacetime fabric, bypassing the physical distance limitations of $v_{ext}$ and achieving instantaneous information transmission.

\subsection{4. Observation \& Reconstruction}

\subsubsection{Measurement-Induced Phase Transition}

The physical mechanism by which observers ``stitch'' reality. By adjusting observation frequency (the intensity of gaze), observers collapse the universe from an ``entangled fluid'' (volume law) into a ``classical crystal'' (area law), thereby establishing the stability of the macroscopic world.

\subsubsection{The Weaver}

A term for awakened observers. Refers to those individuals who realize that ``space is relationship.'' Through actively establishing conscious entanglement (love/understanding), they repair broken warp and weft threads in the fragmented cosmic network, reconstructing the topological structure of reality.

